---
title: "Liveability Index"
output: html_document
date: "2024-08-28"
---

# 1. Diversity Domain Indicators
```{r packages}
# read in packages
library(sf)
library(dplyr)
library(tidyverse)
```

## 1.1. Data Import
```{r read in data}
# read in POIs data
nb_entropy <- st_read('C:/Users/phily/liveability-index/data/pois_london.geojson') %>% 
  select(-c('osmid', 'name','detailed_urban_service', 'lon', 'lat', 
            'GSS_CODE')) %>% 
  st_transform(., 27700)

# read in neighbourhood boundary data
nb_polygon <- st_read('C:/Users/phily/liveability-index/data/neighbourhood/final_neighbourhood_polygons.gpkg') %>%
  st_transform(., 27700) 


# intersect two data
nb_entropy <- st_intersection(nb_entropy, nb_polygon)
```



## 1.2. Entropy Calculation
```{r entropy calculation}
# group by cluster, 15minute service and amenity
diversity <- nb_entropy %>%
  group_by(nb_clus, X15minute_6_urban_service, amenity) %>% 
  summarise(counted = n())


# entropy calculation 
nb_entropy <- diversity %>%
  group_by(nb_clus, X15minute_6_urban_service) %>%
  mutate(total_count = sum(counted)) %>%
  mutate(probability = counted / sum(counted),
         entropy = -sum(probability * log(probability))) %>%
  distinct(nb_clus, X15minute_6_urban_service, .keep_all = TRUE) %>%
  select(nb_clus, X15minute_6_urban_service, entropy)

# stats
min(nb_entropy$entropy) # 0
max(nb_entropy$entropy) # 3.799
```


```{r data format change}
# convert data format to a wide one
# Pivot the table from long to wide format
test <- nb_entropy %>% st_drop_geometry()

nb_entropy_wide <- test %>%
  pivot_wider(names_from = X15minute_6_urban_service, values_from = entropy, names_prefix = "") %>%
  rename(d_c = Commerce, d_l = Living, d_edu = Education,
         d_enter = Entertainment, d_h = Healthcare) 

# fill NA values with -1
# A neighbourhood with no amenities will be assigned a score of -1, while a neighbourhood with only one kind of amenity will get a score of 0 following the entropy equation
nb_entropy_wide <- replace(nb_entropy_wide, is.na(nb_entropy_wide), -1) 
nb_entropy_wide <- nb_entropy_wide %>% select(-Other)


# Save the dataframe as a CSV file
file_path = "C:/Users/phily/liveability-index/data/domain data/diversity/diversity_indicator_nb_450.csv"
st_write(nb_entropy_wide, file_path, row.names = FALSE)

rm(nb_entropy, test)
gc()
```

# 2. Proximity Domain Indicators
```{r packages}
# load packages
library(data.table)
library(doParallel)
library(cppRouting)
library(RcppParallel)
library(concaveman)
```


## 2.1. Origin Points - Centroid of Neighbourhoods
```{r}
# extract centroids of each neighbourhood
nb_centroids <- nb_polygon %>% st_centroid()

# save data
# file_path = "C:/Users/phily/liveability-index/data/domain data/proximity/origin_coords_450.geojson"
# 
# st_write(nb_centroids, file_path, driver = 'GeoJSON')
```

## 2.2. Destination Points - Creating polygons of each urban service
```{r filter centre of mass by number of counts}
# if there are more than 4 points, we create polygons and take the centroid as center of mass 
diversity_poly <- diversity %>% select(-amenity) %>%
  group_by(nb_clus, X15minute_6_urban_service) %>%
  summarise(count = sum(counted)) %>%
  filter(count >= 4)
  

# If there are less than 4 points, you cannot form a polygon - we need to use the mean coordinates
diversity_centroid <- diversity %>% select(-amenity) %>%
  group_by(nb_clus, X15minute_6_urban_service) %>%
  summarise(count = sum(counted)) %>%
  filter(count < 4)
```

### 2.2.1. Center of mass
#### 2.2.1.1. Centre of mass from 1, 2, 3 points
If there is only 1, 2 or 3 amenities in the neighbourhoods,
these points cannot form polygons. That's why the centre of mass is calculated separately.
```{r center of mass for non-polygons}
cm_1 <- diversity_centroid %>% 
  filter(count == 1) %>% select(-count)

cm_2 <- diversity_centroid %>%
  filter(count == 2) %>% st_centroid(geometry) %>% select(-count)

cm_3 <- diversity_centroid %>%
  filter(count == 3) %>% st_centroid(geometry) %>% select(-count)
```


#### 2.2.1.2. center of mass from polygons
If we take the centroids of polygons, we can get the center of mass. This can account for general distribution of amenities
```{r}
# create polygons out of multipoints
polygons <- diversity_poly %>%
  mutate(polygon = st_convex_hull(geometry)) %>% 
  # take out the center of mass - centroid
  mutate(centroid = st_centroid(polygon))

# drop original geometry
cm_4 <- polygons %>%
  select(c(1,2,6)) %>% st_drop_geometry()

# make it as a sf object, centroid as geometry
cm_4 <- cm_4 %>%
  st_as_sf()

cm_4 <- cm_4 %>% st_set_geometry("centroid") %>% rename(geometry = centroid)
  
# rbind
cm_list <- list(cm_1, cm_2, cm_3, cm_4)
dest_coords <- do.call(rbind, cm_list) 
dest_coords <- dest_coords %>%
  group_by(nb_clus, X15minute_6_urban_service)
  

# destination points coordinates - coordinates indicate the center of mass for each urban service within neighbourhoods
dest_coords <- dest_coords %>% arrange(nb_clus)

# file_path = "C:/Users/phily/liveability-index/data/domain data/proximity/dest_coords_450.geojson"
# 
# st_write(dest_coords, file_path, driver = 'GeoJSON')

rm(cm_1, cm_2, cm_3, cm_4, cm_list, diversity, diversity_poly, diversity_centroid, polygons)
gc()
```
All origin and destination points are in the same CRS:27700


## 2.3. Average Euclidean Distance calculation
The average Euclidean distance is calculated here.
Origin point is the centroid of neighbourhoods 
Destination point is the center of mass for each urban service within the neighbourhoods.

This approach captures the average distance from the neighborhood centre to the urban service-related amenities as a whole.

By using center of mass as destination, we can effectively consider the overall distribution of urban services and take into account their collective proximity to the center of neighbourhoods.

```{r proximity calculation}
# Create an empty vector to store the distances
distances <- vector("numeric", nrow(dest_coords))

# Iterate over the rows of the destination data
for (i in 1:nrow(dest_coords)) {
  # Get the nb_clus value for the current destination point
  nb_clus <- dest_coords$nb_clus[i]
  
  # Find the matching origin point based on nb_clus
  origin <- nb_centroids[nb_centroids$nb_clus == nb_clus, ]
  
  # Calculate the Euclidean distance between the origin and destination points
  distance <- st_distance(origin, dest_coords[i, ])
  
  # Store the distance in the vector
  distances[i] <- distance
}

# Add the distances as a new column to the destination sf object
# distance here indicates the distance between the centroid of neighbourhoods to
# center of mass of urban services within neighbourhoods
dest_coords$distance <- distances

# max?
max(dest_coords$distance) # 1657

# min?
min(dest_coords$distance) # 1.763

rm(origin, distance, distances)
gc()
```


```{r save data}
# save data
proximity <- dest_coords %>% st_drop_geometry()

proximity_wide <- proximity %>%
  pivot_wider(names_from = X15minute_6_urban_service, 
              values_from = distance, names_prefix = "") %>%
  rename(p_c = Commerce, p_l = Living, p_edu = Education,
         p_enter = Entertainment, p_h = Healthcare)


# As there are NA values, an additional column has been created. remove it
proximity_wide <- proximity_wide %>% select(-Other)

rm(proximity)
gc()
```


## 2.4. Network Construction
For those neighbourhoods which miss any of the five urban services, the network distance to the closest center of mass for that missing urban service in neighbouring neighbourhoods will be calculated.

```{r}
edges <- st_read('C:/Users/phily/liveability-index/data/network/sf_all_london.gpkg', layer = "edges") %>% st_transform(., 27700)

# create unique edge identifier
edges <- edges %>% dplyr::select(from, to, highway, length, geom) %>% dplyr::mutate(edgeID = c(1:n())) %>%
  dplyr::select(edgeID, from, to, highway, length, geom)


# create nodes at the start and end point of each edge
# give each node a unique index
nodes <- edges %>% 
  st_coordinates() %>%
  as_tibble() %>%
  dplyr::rename(edgeID = L1) %>%
  group_by(edgeID) %>%
  slice(c(1,n())) %>%
  ungroup() %>%
  dplyr::mutate(start_end = rep(c('start', 'end'), times = n()/2)) %>%
  dplyr::mutate(xy = paste(.$X, .$Y)) %>%
  
  # by using factor(), we can see many unique point geometry
  # based on this, we give node indices
  dplyr::mutate(nodeID = group_indices(., factor(xy, levels = unique(xy)))) %>%
  dplyr::select(-xy)


# start, end points for each edge
# pull out start nodes
start_nodes <- nodes %>% 
  filter(start_end == 'start') %>%
  pull(nodeID)

# pull out end nodes
end_nodes <- nodes %>%
  filter(start_end == 'end') %>%
  pull(nodeID)

# plug in start/end nodes and reorder columns
edges <- edges %>% select(-c(from, to)) %>%
  mutate(from = start_nodes, to = end_nodes) %>%
  select(from, to, highway, length, edgeID, geom)

# remove end_nodes and start_nodes to save memory usage
rm(end_nodes)
rm(start_nodes)

# remove duplicate nodes & set the same crs (EPSG:27700)
nodes <- nodes %>% 
  # keep unique node ID
  distinct(nodeID, .keep_all = TRUE) %>%
  dplyr::select(-c(edgeID, start_end)) 
```


```{r create network}
# set the number of threads used by cppRouting
RcppParallel::setThreadOptions(numThreads = 1)

# prepare data to an appropriate format
roads <- edges %>% select(from, to, length) %>% st_drop_geometry() %>% as.data.table()

coord <- nodes %>% select(nodeID, X, Y) %>% as.data.table()

# instantiate a graph with coordinates
graph  <-  roads %>% 
  makegraph(directed = F,
            coords = coord)

# remove edges and nodes to save memory
rm(edges)
rm(nodes)
rm(coord)
rm(roads)
```


```{r simplify graph}
# remove non-intersection nodes, duplicated edges and isolated loops - simplify graph
graph <- graph %>%
  cppRouting::cpp_simplify(rm_loop = T,
                           iterate = F)
```


```{r finding nearest node on the graph function}
# this function finds the nearest nodes on the graph
find_nearest_nodes_on_graph_sf <- function(point_data, graph, crs = 27700) {
  
  nodes_on_graph <- graph$coords %>% st_as_sf(., 
                                              coords = c('X', 'Y'),
                                              crs = 27700) 
  
  poi_data <- point_data %>% st_transform(27700)
  
  graph$coords[st_nearest_feature(poi_data, nodes_on_graph),'nodeID']
  
  
}
```

## 2.5. Network Distance
### 2.5.1. p_edu calculation using network distance
The proximity to urban services of the neighbourhoods which miss any of five urban services will be measured between the center of neighbourhoods to the nearest centre of mass in other neighbourhoods, using network distance
```{r calculate distance for those neighbourhoods with missing education services}
# P_edu - Education
# pull out the nb_clus which has education services - they will be destination points 396 neighbourhoods
edu_dest_nb_clus <- proximity_wide %>% filter(!is.na(p_edu)) %>% pull(nb_clus)

# Filter the nb_centroids data frame using the extracted nb_clus values
edu_dest <- dest_coords %>% 
  filter(nb_clus %in% edu_dest_nb_clus) %>% st_as_sf() %>% 
  filter(X15minute_6_urban_service == 'Education') #%>% st_cast('POINT')

# pull the nb_clus which does not have education 
edu_nb_clus <- proximity_wide %>% filter(is.na(p_edu)) %>% pull(nb_clus) 

# Filter the nb_centroids data frame using the extracted nb_clus values
edu_ori <- nb_centroids %>% 
  filter(nb_clus %in% edu_nb_clus) %>% st_as_sf()

rm(edu_dest_nb_clus, edu_nb_clus, i, nb_clus)
gc()
```


This projects the nodes on the graph and finds the closest corresponding nodes on the graph
```{r find closest nodes in the graph}
nearest_nodes_origin <- find_nearest_nodes_on_graph_sf(edu_ori, graph, crs = 27700)
nearest_nodes_dest <- find_nearest_nodes_on_graph_sf(edu_dest, graph, crs = 27700)
```



```{r calculate network distance}
dist <- get_distance_matrix(graph,
                            from = nearest_nodes_origin$nodeID,
                            to = nearest_nodes_dest$nodeID)

# extract minimum travel from each node
dist2<-data.frame(node=nearest_nodes_origin$nodeID,mindist=apply(dist,1,min,na.rm=T))

# joining edu_ori nb_clus to nodeID
edu_ori$nodeID <- nearest_nodes_origin %>% pull(nodeID)
edu_ori_1 <- edu_ori %>% left_join(.,
                                 dist2,
                                 by = c('nodeID' = 'node')) 


# embed this distance value to original proximity_wide dataframe
edu_ori_test <- edu_ori_1 %>% st_drop_geometry() %>% select(nb_clus, mindist)
test <- left_join(proximity_wide,
                  edu_ori_test,
                  by = 'nb_clus')

test$p_edu <- ifelse(is.na(test$p_edu), test$mindist, test$p_edu)

# Drop the mindist column, as it is no longer needed
proximity_wide <- test %>% select(-mindist)

rm(edu_dest, edu_ori, edu_ori_1, edu_ori_test)
gc()
```



### 2.5.2. p_l calculation using network distance
All neighbourhoods have living facilities.
```{r calculate distance for those neighbourhoods with missing urban services}
# P_l - Living
# pull out the nb_clus which has living facilities - 399 destinations
living_dest_nb_clus <- proximity_wide %>% filter(!is.na(p_l)) %>% pull(nb_clus)

```


### 2.5.3. p_c calculation using network distance
```{r calculate distance for those neighbourhoods with missing urban services}
# P_c - Commerce
# pull out the nb_clus which has commerce - 396 facilities
com_dest_nb_clus <- proximity_wide %>% filter(!is.na(p_c)) %>% pull(nb_clus)

# Filter the nb_centroids data frame using the extracted nb_clus values
com_dest <- dest_coords %>% 
  filter(nb_clus %in% com_dest_nb_clus) %>% st_as_sf() %>% 
  filter(X15minute_6_urban_service == 'Commerce') #%>% st_cast('POINT')

# pull the nb_clus which does not have commerce - 3 origin 
com_nb_clus <- proximity_wide %>% filter(is.na(p_c)) %>% pull(nb_clus) 

# Filter the nb_centroids data frame using the extracted nb_clus values
com_ori <- nb_centroids %>% 
  filter(nb_clus %in% com_nb_clus) %>% st_as_sf()

sum(is.na(proximity_wide$p_c)) # 3

rm(com_dest_nb_clus, com_nb_clus)
gc()
```

```{r find closest nodes in the graph}
nearest_nodes_origin <- find_nearest_nodes_on_graph_sf(com_ori, graph, crs = 27700)
nearest_nodes_dest <- find_nearest_nodes_on_graph_sf(com_dest, graph, crs = 27700)
```


```{r calculate network distance}
dist <- get_distance_matrix(graph,
                            from = nearest_nodes_origin$nodeID,
                            to = nearest_nodes_dest$nodeID)

# extract minimum travel
dist2<-data.frame(node=nearest_nodes_origin$nodeID,mindist=apply(dist,1,min,na.rm=T))

# joining edu_ori nb_clus to nodeID
com_ori$nodeID <- nearest_nodes_origin %>% pull(nodeID)
com_ori_1 <- com_ori %>% left_join(.,
                                 dist2,
                                 by = c('nodeID' = 'node')) 


# imbed this distance value to original proximity_wide dataframe
com_ori_test <- com_ori_1 %>% st_drop_geometry() %>% select(nb_clus, mindist)
test <- left_join(proximity_wide,
                  com_ori_test,
                  by = 'nb_clus')

test$p_c <- ifelse(is.na(test$p_c), test$mindist, test$p_c)

# Drop the mindist column, as it is no longer needed
proximity_wide <- test %>% select(-mindist)

sum(is.na(proximity_wide$p_c))

rm(com_dest, com_ori, com_ori_1, com_ori_test)
gc()
```



### 2.5.4. p_enter calculation using network distance
```{r calculate distance for those neighbourhoods with missing urban services}
# P_enter - Entertainment
# pull out the nb_clus which has entertainment - 397
enter_dest_nb_clus <- proximity_wide %>% filter(!is.na(p_enter)) %>% pull(nb_clus)

# Filter the nb_centroids data frame using the extracted nb_clus values
enter_dest <- dest_coords %>% 
  filter(nb_clus %in% enter_dest_nb_clus) %>% st_as_sf() %>% 
  filter(X15minute_6_urban_service == 'Entertainment') #%>% st_cast('POINT')

# pull the nb_clus which does not have entertainment - 2
enter_nb_clus <- proximity_wide %>% filter(is.na(p_enter)) %>% pull(nb_clus) 

# Filter the nb_centroids data frame using the extracted nb_clus values
enter_ori <- nb_centroids %>% 
  filter(nb_clus %in% enter_nb_clus) %>% st_as_sf()

sum(is.na(proximity_wide$p_enter)) # 2

rm(enter_dest_nb_clus, enter_nb_clus)
gc()
```


```{r find closest nodes in the graph}
nearest_nodes_origin <- find_nearest_nodes_on_graph_sf(enter_ori, graph, crs = 27700)
nearest_nodes_dest <- find_nearest_nodes_on_graph_sf(enter_dest, graph, crs = 27700)
```


```{r calculate network distance}
dist <- get_distance_matrix(graph,
                            from = nearest_nodes_origin$nodeID,
                            to = nearest_nodes_dest$nodeID)

# extract minimum travel
dist2<-data.frame(node=nearest_nodes_origin$nodeID,mindist=apply(dist,1,min,na.rm=T))

# joining enter_ori nb_clus to nodeID
enter_ori$nodeID <- nearest_nodes_origin %>% pull(nodeID)
enter_ori_1 <- enter_ori %>% left_join(.,
                                 dist2,
                                 by = c('nodeID' = 'node')) 


# imbed this distance value to original proximity_wide dataframe
enter_ori_test <- enter_ori_1 %>% st_drop_geometry() %>% select(nb_clus, mindist)
test <- left_join(proximity_wide,
                  enter_ori_test,
                  by = 'nb_clus')

test$p_enter <- ifelse(is.na(test$p_enter), test$mindist, test$p_enter)

# Drop the mindist column, as it is no longer needed
proximity_wide <- test %>% select(-mindist)

sum(is.na(proximity_wide$p_enter))

rm(enter_dest, enter_ori, enter_ori_1, enter_ori_test)
gc()
```



### 2.5.5. p_h calculation using network distance
```{r calculate distance for those neighbourhoods with missing urban services}
# P_h - Healthcare
# pull out the nb_clus which has health - 384
health_dest_nb_clus <- proximity_wide %>% filter(!is.na(p_h)) %>% pull(nb_clus)

# Filter the nb_centroids data frame using the extracted nb_clus values
health_dest <- dest_coords %>% 
  filter(nb_clus %in% health_dest_nb_clus) %>% st_as_sf() %>% 
  filter(X15minute_6_urban_service == 'Healthcare') #%>% st_cast('POINT')

# pull the nb_clus which does not have health - 15
health_nb_clus <- proximity_wide %>% filter(is.na(p_h)) %>% pull(nb_clus) 

# Filter the nb_centroids data frame using the extracted nb_clus values
health_ori <- nb_centroids %>% 
  filter(nb_clus %in% health_nb_clus) %>% st_as_sf()

sum(is.na(proximity_wide$p_h)) # 15

rm(health_dest_nb_clus, health_nb_clus)
gc()
```


```{r find closest nodes in the graph}
nearest_nodes_origin <- find_nearest_nodes_on_graph_sf(health_ori, graph, crs = 27700)
nearest_nodes_dest <- find_nearest_nodes_on_graph_sf(health_dest, graph, crs = 27700)
```


```{r calculate network distance}
dist <- get_distance_matrix(graph,
                            from = nearest_nodes_origin$nodeID,
                            to = nearest_nodes_dest$nodeID)

# extract minimum travel
dist2<-data.frame(node=nearest_nodes_origin$nodeID,mindist=apply(dist,1,min,na.rm=T))

# joining enter_ori nb_clus to nodeID
health_ori$nodeID <- nearest_nodes_origin %>% pull(nodeID)
health_ori_1 <- health_ori %>% left_join(.,
                                 dist2,
                                 by = c('nodeID' = 'node')) 


# embed this distance value to original proximity_wide dataframe
health_ori_test <- health_ori_1 %>% st_drop_geometry() %>% select(nb_clus, mindist)
test <- left_join(proximity_wide,
                  health_ori_test,
                  by = 'nb_clus')

test$p_h <- ifelse(is.na(test$p_h), test$mindist, test$p_h)
# Drop the mindist column, as it is no longer needed
proximity_wide <- test %>% select(-mindist)

sum(is.na(proximity_wide$p_h)) #0

max(proximity_wide$p_edu) # 2635
max(proximity_wide$p_c) # 3414
max(proximity_wide$p_l) # 874
max(proximity_wide$p_enter) # 2153
max(proximity_wide$p_h) # 3070

rm(health_dest, health_ori, health_ori_1, health_ori_test, find_nearest_nodes_on_graph_sf, test, nearest_nodes_dest, nearest_nodes_origin, dist, dist2, dest_coords, graph, nb_centroids)
gc()
```



```{r save proximity data}
file_path = "C:/Users/phily/liveability-index/data/domain data/proximity/proximity_indicator_nb_450.csv"

write.csv(proximity_wide, file_path, row.names = FALSE)
```


# 3. Population Density Indicator
LSOA-level populationd data was projected onto hexagon grids at Uber H3 level 10, providing a consistent basis for aggregation. Then, it was projected onto neighbourhoods. 
```{r read in hex-level pop data}
# read in hexagon-level population data
hex_pop <- st_read('C:/Users/phily/liveability-index/data/domain data/pop/hexagon_pop.geojson') %>% st_transform(27700)
```


```{r cumulative population at neighbourhood level}
# calculate the area size of each neighbourhood
hex_pop$hex_area <- st_area(hex_pop)

# Perform spatial intersection
intersection <- st_intersection(hex_pop, nb_polygon)

# Calculate the proportion of each hexagon's area within each neighborhood polygon
intersection$area_proportion <- st_area(intersection) / intersection$hex_area

# Calculate the estimated population within each neighborhood polygon
intersection$estimated_population <- intersection$weighted_avg_pop * intersection$area_proportion

intersection <- intersection %>%
  group_by(nb_clus) %>%
  summarise(populations = sum(estimated_population)) 

# calculate area size
intersection <- intersection %>% mutate(area = st_area(.))

# Calculate population density per square kilometer
intersection <- intersection %>%
  mutate(pop_density = populations / (area / 1000000))
                                      
max(intersection$pop_density)
min(intersection$pop_density)

# select columns
intersection <- intersection %>% st_drop_geometry()
intersection <- intersection %>% select(c(1,4))

# save data - just those with populations
# file_path = 'C:/Users/phily/liveability-index/data/domain data/pop/pop_density_nb_450.csv'
# 
# st_write(intersection, file_path)

rm(hex_pop)
gc()
```


# 4. Data Consolidation
```{r merge data}
# merge diversity and proximity indicators
div_prox_450 <- nb_entropy_wide %>% left_join(.,
                                            proximity_wide,
                                            by = 'nb_clus')

final_450 <- div_prox_450 %>% left_join(.,
                                        intersection,
                                        by = 'nb_clus')

# save data
file_path = 'C:/Users/phily/liveability-index/data/domain data/raw_domain_indicator_nb_450.csv'
write.csv(final_450, file_path, row.names = FALSE)

rm(div_prox_450, intersection, proximity_wide, nb_entropy_wide)
gc()
```


# 5. Normalisation and Aggregation
Normalisation was done using exponential transformation, which was used in creating Indices of Multiple Deprevation by UK government. This normalisation first ranks each area by its score and then transforms into a scale ranging between 1 and 100.
```{r}
# read in data
# df <- read.csv('C:/Users/phily/liveability-index/data/domain data/raw_domain_indicator_nb_450.csv')

df <- final_450

diversity <- df[1:6]
proximity <- df[c(1, 7:11)]
pop_density <- df[c(1, 12)]

rm(final_450)
gc()
```

## 5.1. Diversity Domain
### 5.1.1. Custom Functions for Diversity Domain
```{r functions for diversity}
# exponential transformation function
# The higher the value is, the higher transformed value it gets.
# As higher value indicates the better diversity of neighbourhood
exp_trans <- function(var){
  -23 * log(1 - (var * (1 - (exp(-100/23)))))
}

# function to rank and exponentially transform a variable in a given table
rank_exp <- function(df, var){
  var <- enquo(var)

  df %>%
    ungroup() %>%   # <- critical
    mutate(rank   = min_rank(!! var),
           rank_s = scales::rescale(rank, to = c(0, 1)),
           exp    = exp_trans(rank_s)) %>%
    pull(exp)
}

```



### 5.1.2. Diversity Domain Normalisation
```{r diversity}
# exponential transformation on each diversity indicator
diversity[,'d_c_exp'] <- rank_exp(diversity, d_c)
diversity[,'d_edu_exp'] <- rank_exp(diversity, d_edu)
diversity[,'d_enter_exp'] <- rank_exp(diversity, d_enter)
diversity[,'d_l_exp'] <- rank_exp(diversity, d_l)
diversity[,'d_h_exp'] <- rank_exp(diversity, d_h)

# diversity domain score
diversity <- diversity %>%
  mutate(diversity_score = (d_c_exp / 5) + (d_edu_exp / 5) +
           (d_enter_exp / 5) + (d_l_exp / 5) + (d_h_exp / 5))

# Overall diversity exponential transformation
diversity['diversity_exp'] <- rank_exp(diversity, diversity_score)

# final diversity data
diversity_final <- diversity %>% select(nb_clus, d_c_exp, d_edu_exp,
                                        d_enter_exp, d_l_exp,
                                        d_h_exp, diversity_score,
                                        diversity_exp)
```

## 5.2. Proximity Domain
### 5.2.1. Custom Functions for Proximity Domain
```{r}
# exponential transformation function
# Unlike diversity, the shorter the distance is, the better the proximity is.
# So, we have to modify the previous function.
# We have to reverse the rank as this is opposite of diversity
# Exponential transformation function
exp_trans <- function(var){
  -23 * log(1 - (var * (1 - (exp(-100/23)))))
}

# Function to rank and exponentially transform a variable in a given table
rank_exp_proximity <- function(df, var){
  var <- enquo(var)
  
  df %>%
    ungroup() %>%  # remove any grouping
    mutate(
      rank = min_rank(!!var),
      reversed_rank = max(rank) - rank + 1,          # reverse rank for proximity
      rank_s = scales::rescale(reversed_rank, to = c(0, 1)), # rescale to 0-1
      exp = exp_trans(rank_s)
    ) %>%
    pull(exp)  # return as a numeric vector
}

```

### 5.2.2. Proximity Domain Normalisation
```{r proximity }
# exponential transformation on each proximity indicator
proximity[,'p_c_exp'] <- rank_exp_proximity(proximity, p_c)
proximity[,'p_enter_exp'] <- rank_exp_proximity(proximity, p_enter)
proximity[,'p_edu_exp'] <- rank_exp_proximity(proximity, p_edu)
proximity[,'p_l_exp'] <- rank_exp_proximity(proximity, p_l)
proximity[,'p_h_exp'] <- rank_exp_proximity(proximity, p_h)

# proximity score
proximity <- proximity %>%
  mutate(proximity_score = (p_c_exp / 5) + (p_edu_exp / 5) +
           (p_enter_exp / 5) + (p_l_exp / 5) + (p_h_exp / 5))

# Overall proximity exponential transformation
# Attention: When calculating overall proximity exponential transformation,
# the higher proximity score, which is combined by 5 proximity subindicators, should receive a higher rank
# which is different from calculating the proximity subindicators where we gave the lower distance a higher # exponential rank.
# So, here we use the original 'rank_exp' function to credit those neighbourhoods with high proximity score with higher ranks.
# If we use the function (rank_exp_proximity) used for calculating proximity subindicators, we will commit an error of giving high proximity scores a lower rank.
proximity['proximity_exp'] <- rank_exp(proximity, proximity_score)

# final proximity data
proximity_final <- proximity %>% select(nb_clus, p_c_exp, p_edu_exp,
                                        p_enter_exp, p_l_exp,
                                        p_h_exp, proximity_score,
                                        proximity_exp)
```


## 5.3. Population Density
```{r}
pop_density['pop_density_exp'] <- rank_exp(pop_density, pop_density)

pop_density_final <- pop_density %>% select(nb_clus,
                                            pop_density_exp)
```

## 5.4. Final Index
```{r}
# left join diversity and proximity domain
li_d_p <- diversity_final %>%
  left_join(proximity_final,
            by = 'nb_clus')

# left join previous data with population density domain
li_d_p_pop <- li_d_p %>%
  left_join(pop_density_final,
            by = 'nb_clus')

# liveability index - combine three domains
li <- li_d_p_pop %>%
  mutate(li_score = (diversity_exp / 3) + (proximity_exp / 3) + 
           (pop_density_exp / 3))

# calculate the total score
li['li_exp'] <- rank_exp(li, li_score)

# calculate decile
li_final <- li %>%
  ungroup() %>%   # remove any grouping
  mutate(
    li_rank = min_rank(li_exp),
    li_decile = ntile(li_exp, 10)
  )


# save data
file_path = 'C:/Users/phily/liveability-index/data/domain data/final_normalised_domain_indicator_nb_450.csv'
# 
write.csv(li_final, file_path, row.names = FALSE)

rm(exp_trans, rank_exp, rank_exp_proximity, proximity, proximity_final, diversity, diversity_final, li, li_d_p, li_d_p_pop, pop_density, pop_density_final)
gc()
```

## 5.5. PCA
```{r}
# read in data
df <- read.csv('C:/Users/phily/liveability-index/data/domain data/raw_domain_indicator_nb_450.csv')
```


```{r}
# extract diversity indicators
diversity <- df[2:6]

proximity <- df[7:11]


# correlation
cor(diversity)
cor(proximity)
```

### 5.5.1. Diversity PCA
```{r}
plot(diversity)
```

```{r diversity}
# PCA
pc <- diversity %>% prcomp(center = T, scale = T)
summary(pc)
```
First four PCs explain more than 90% of total variance

```{r loadings}
print(pc)
```
The loadings represent the correlation between each variable and the corresponding principal component. The higher the absolute value of the loading, the more influence the variable has on that PC.
Each row represents a variable, and the values in each column represent the loadings of that variable in the respective principal component. For instance, the loadings of d_l in PC2 is approximately 0.941, indicating a positive correlation.

It seems that PC2 is related to d_l, PC3 is related to d_edu (0.882), PC4 is related to d_enter (0.778), PC5 is related to d_c (0.742) and d_h (0.665). This indicates that the variables have unique and distinct patterns of variability. And grouping them into few PCs may lead to loss of information and the final composite index might not best present each indicator efficiently.

```{r}
# Access the eigenvalues of each PC from the PCA object
eigenvalues <- pc$sdev^2

# Print the eigenvalues
print(eigenvalues)
```

### 5.5.2. Proxmity PCA
```{r proximity}
# PCA
pc_prox <- proximity %>% prcomp(center = T, scale = T)
summary(pc_prox)
```

First four components explain more than 90% of the total variance.

```{r proximity loadings}
print(pc_prox)
```
p_l is highly related to PC4 (0.747), p_enter is related to PC2 (0.710), p_c is related to PC4 (0.651), p_h is related to PC5 (0.713), and p_edu is correlated to PC3 (0.672).

```{r}
# Access the eigenvalues of each PC from the PCA object
eigenvalues <- pc_prox$sdev^2

# Print the eigenvalues
print(eigenvalues)
```



# 6. Uncertainty Test
```{r packages}
library(tmap)
library(tmaptools)
```
## 6.1. Prerequisites
### 6.1.1. Customised functions
```{r custom functions}
# exponential transformation
exp_trans <- function(var){
  -23 * log(1 - (var * (1 - (exp(-100/23)))))
}


# function to rank and exponentially transform a variable in a given table
rank_exp <- function(df, var){
  var <- enquo(var)

  df %>%
    ungroup() %>%   # <- critical
    mutate(rank   = min_rank(!! var),
           rank_s = scales::rescale(rank, to = c(0, 1)),
           exp    = exp_trans(rank_s)) %>%
    pull(exp)
}


# proximity 
rank_exp_proximity <- function(df, var){
  var <- enquo(var)
  
  df %>%
    ungroup() %>%  # remove any grouping
    mutate(
      rank = min_rank(!!var),
      reversed_rank = max(rank) - rank + 1,          # reverse rank for proximity
      rank_s = scales::rescale(reversed_rank, to = c(0, 1)), # rescale to 0-1
      exp = exp_trans(rank_s)
    ) %>%
    pull(exp)  # return as a numeric vector
}


# minmax scalar
minmax <- function(df, var, na.rm = TRUE) {
  var <- enquo(var)
  
  df %>%
    ungroup() %>%
    mutate(
      # convert to numeric to ignore units
      mm = (as.numeric(!!var) - min(as.numeric(!!var), na.rm = na.rm)) /
           (max(as.numeric(!!var), na.rm = na.rm) - min(as.numeric(!!var), na.rm = na.rm))
    ) %>%
    pull(mm)
}

minmax_proximity <- function(df, var, na.rm = TRUE) {
  var <- enquo(var)
  
  df %>%
    ungroup() %>%
    mutate(
      mm = 1 - ((as.numeric(!!var) - min(as.numeric(!!var), na.rm = na.rm)) /
                (max(as.numeric(!!var), na.rm = na.rm) - min(as.numeric(!!var), na.rm = na.rm)))
    ) %>%
    pull(mm)
}


# Z-score standardisation
ss <- function(df, var, na.rm = TRUE) {
  var <- enquo(var)
  
  df %>%
    ungroup() %>%
    mutate(
      ss = (as.numeric(!!var) - mean(as.numeric(!!var), na.rm = na.rm)) /
           sd(as.numeric(!!var), na.rm = na.rm)
    ) %>%
    pull(ss)
}

ss_proximity <- function(df, var, na.rm = TRUE) {
  var <- enquo(var)
  
  df %>%
    ungroup() %>%
    mutate(
      ss = -((as.numeric(!!var) - mean(as.numeric(!!var), na.rm = na.rm)) /
             sd(as.numeric(!!var), na.rm = na.rm))
    ) %>%
    pull(ss)
}

# Robust normalisation
robust <- function(df, var, na.rm = TRUE) {
  var <- enquo(var)
  
  df %>%
    ungroup() %>%
    mutate(
      rs = (as.numeric(!!var) - median(as.numeric(!!var), na.rm = na.rm)) /
           IQR(as.numeric(!!var), na.rm = na.rm)
    ) %>%
    pull(rs)
}

robust_proximity <- function(df, var, na.rm = TRUE) {
  var <- enquo(var)
  
  df %>%
    ungroup() %>%
    mutate(
      rs = -((as.numeric(!!var) - median(as.numeric(!!var), na.rm = na.rm)) /
              IQR(as.numeric(!!var), na.rm = na.rm))
    ) %>%
    pull(rs)
}


```

### 6.1.2. Lists to use in a loop
We have 288 scenarios (6 * 6 * 2 * 4)
```{r}
# Exclusion of subindicators
## weights given, when one of subindicators is excluded
diversity_weights <- list('1' = c(0.2, 0.2, 0.2, 0.2, 0.2),
                          '2' = c(0, 0.25, 0.25, 0.25, 0.25),
                          '3' = c(0.25, 0, 0.25, 0.25, 0.25),
                          '4' = c(0.25, 0.25, 0, 0.25, 0.25),
                          '5' = c(0.25, 0.25, 0.25, 0, 0.25),
                          '6' = c(0.25, 0.25, 0.25, 0.25, 0))

proximity_weights <- list('1' = c(0.2, 0.2, 0.2, 0.2, 0.2),
                          '2' = c(0, 0.25, 0.25, 0.25, 0.25),
                          '3' = c(0.25, 0, 0.25, 0.25, 0.25),
                          '4' = c(0.25, 0.25, 0, 0.25, 0.25),
                          '5' = c(0.25, 0.25, 0.25, 0, 0.25),
                          '6' = c(0.25, 0.25, 0.25, 0.25, 0))

# When combining 3 domains, in case when population density is excluded
pop_density_weights <- list('1' = c(1/3, 1/3, 1/3),
                            '2' = c(0.5, 0.5, 0))

# normalisation
norm_weights <- list('1' = c(rank_exp, rank_exp_proximity),
                     '2' = c(minmax, minmax_proximity),
                     '3' = c(ss, ss_proximity),
                     '4' = c(robust, robust_proximity))

```


## 6.2. Processing
### 6.2.1. Rank Exponential Normalisation
```{r 72 scenarios - rank_exp}
ua_output_exp <- tibble()

# run for loop - creates 72 versions of LI and store results for each neighbourhood

for (div_w in seq(1, length(diversity_weights))) {
  
  # diversity subindicators
  diversity <- df[1:6]
  
  # exponential transformation on each diversity subindicator
  diversity[,'d_c_exp'] <- rank_exp(diversity, d_c)
  diversity[,'d_edu_exp'] <- rank_exp(diversity, d_edu)
  diversity[,'d_enter_exp'] <- rank_exp(diversity, d_enter)
  diversity[,'d_l_exp'] <- rank_exp(diversity, d_l)
  diversity[,'d_h_exp'] <- rank_exp(diversity, d_h)
  
  # diversity domain score
  div_weight <- diversity %>%
    mutate(div_score = (d_c_exp * diversity_weights[[div_w]][1]) + 
             (d_edu_exp * diversity_weights[[div_w]][2]) + 
             (d_enter_exp * diversity_weights[[div_w]][3]) + 
             (d_l_exp * diversity_weights[[div_w]][4]) + 
             (d_h_exp * diversity_weights[[div_w]][5]))
  
  # diversity final weighting
  div_weight['div_exp'] <- rank_exp(div_weight, div_score)
  div_fin <- select(div_weight, nb_clus, div_exp)
  
  
  
  for (prox_w in seq(1, length(proximity_weights))) {
    
    # proximity subindicator
    proximity <- df[c(1, 7:11)]
    
    # exponential transformation on each proximity indicator
    proximity[,'p_c_exp'] <- rank_exp_proximity(proximity, p_c)
    proximity[,'p_edu_exp'] <- rank_exp_proximity(proximity, p_edu)
    proximity[,'p_enter_exp'] <- rank_exp_proximity(proximity, p_enter)
    proximity[,'p_l_exp'] <- rank_exp_proximity(proximity, p_l)
    proximity[,'p_h_exp'] <- rank_exp_proximity(proximity, p_h)
    
    # proximity domain score
    prox_weight <- proximity %>%
      mutate(prox_score = (p_c_exp * proximity_weights[[prox_w]][1]) + 
               (p_edu_exp * proximity_weights[[prox_w]][2]) + 
               (p_enter_exp * proximity_weights[[prox_w]][3]) + 
               (p_l_exp * proximity_weights[[prox_w]][4]) + 
               (p_h_exp * proximity_weights[[prox_w]][5]))
    
    # proximity final weighting
    prox_weight['prox_exp'] <- rank_exp(prox_weight, prox_score)
    prox_fin <- select(prox_weight, nb_clus, prox_exp)
    
    
    
    for (pop_w in seq(1, length(pop_density_weights))) {
      
      # population density
      pop_density <- df[c(1, 12)]
      
      # exponential transformation on population density
      pop_density['pop_den_exp'] <- rank_exp(pop_density, pop_density)
      
      
      # combine three domains
      div_prox <- left_join(div_fin,
                            prox_fin,
                            by = 'nb_clus')
      
      div_prox_pop <- left_join(div_prox,
                                pop_density,
                                by = 'nb_clus')
      
      # final score
      li_weight <- div_prox_pop %>% 
        mutate(li_score = (div_exp * pop_density_weights[[pop_w]][1]) + 
                 (prox_exp * pop_density_weights[[pop_w]][2]) + 
                 (pop_den_exp * pop_density_weights[[pop_w]][3]))
      
      # liveability score final weighting
      li_weight['li_exp'] <- rank_exp(li_weight, li_score)
      
      # rank and decile
      li_weight <- li_weight %>% ungroup() %>% mutate(li_rank = min_rank(li_exp),
                                      li_decile = ntile(li_exp, 10))
      

      # print current iteration
      print(paste('div_w: ', div_w, ' prox_w: ', prox_w, 'pop_w: ', pop_w))
      
      # create a final output table
      li_final <- select(li_weight, nb_clus, li_score, li_exp, li_rank, li_decile)
      
      # add variable fields
      li_final$diversity_weight <- div_w
      li_final$proximity_weight <- prox_w
      li_final$pop_density_weight <- pop_w
      li_final$normalisation <- 'rank_exp'
      
      ua_output_exp <- rbind(ua_output_exp, li_final)
      
      
    }
    
  }
}

```


### 6.2.2. MinMax Normalisation
```{r 72 scenarios - minmax}
ua_output_mm <- tibble()

# run for loop - creates 72 versions of LI and store results for each neighbourhood

for (div_w in seq(1, length(diversity_weights))) {
  
  # diversity subindicators
  diversity <- df[1:6]
  
  # minmax normalisation on each diversity subindicator
  diversity[,'d_c_mm'] <- minmax(diversity, d_c)
  diversity[,'d_edu_mm'] <- minmax(diversity, d_edu)
  diversity[,'d_enter_mm'] <- minmax(diversity, d_enter)
  diversity[,'d_l_mm'] <- minmax(diversity, d_l)
  diversity[,'d_h_mm'] <- minmax(diversity, d_h)
  
  # diversity domain score
  div_weight <- diversity %>%
    mutate(div_score = (d_c_mm * diversity_weights[[div_w]][1]) + 
             (d_edu_mm * diversity_weights[[div_w]][2]) + 
             (d_enter_mm * diversity_weights[[div_w]][3]) + 
             (d_l_mm * diversity_weights[[div_w]][4]) + 
             (d_h_mm * diversity_weights[[div_w]][5]))
  
  # diversity final weighting
  div_weight['div_mm'] <- minmax(div_weight, div_score)
  div_fin <- select(div_weight, nb_clus, div_mm)
  
  
  
  for (prox_w in seq(1, length(proximity_weights))) {
    
    # proximity subindicator
    proximity <- df[c(1, 7:11)]
    
    # minmax normalisation on each proximity indicator
    proximity[,'p_c_mm'] <- minmax_proximity(proximity, p_c)
    proximity[,'p_edu_mm'] <- minmax_proximity(proximity, p_edu)
    proximity[,'p_enter_mm'] <- minmax_proximity(proximity, p_enter)
    proximity[,'p_l_mm'] <- minmax_proximity(proximity, p_l)
    proximity[,'p_h_mm'] <- minmax_proximity(proximity, p_h)
    
    # proximity domain score
    prox_weight <- proximity %>%
      mutate(prox_score = (p_c_mm * proximity_weights[[prox_w]][1]) + 
               (p_edu_mm * proximity_weights[[prox_w]][2]) + 
               (p_enter_mm * proximity_weights[[prox_w]][3]) + 
               (p_l_mm * proximity_weights[[prox_w]][4]) + 
               (p_h_mm * proximity_weights[[prox_w]][5]))
    
    # proximity final weighting
    prox_weight['prox_mm'] <- minmax(prox_weight, prox_score)
    prox_fin <- select(prox_weight, nb_clus, prox_mm)
    
    
    
    for (pop_w in seq(1, length(pop_density_weights))) {
      
      # population density
      pop_density <- df[c(1, 12)]
      
      # minmax normalisation on population density
      pop_density['pop_den_mm'] <- minmax(pop_density, pop_density)
      
      
      # combine three domains
      div_prox <- left_join(div_fin,
                            prox_fin,
                            by = 'nb_clus')
      
      div_prox_pop <- left_join(div_prox,
                                pop_density,
                                by = 'nb_clus')
      
      # final score
      li_weight <- div_prox_pop %>% 
        mutate(li_score = (div_mm * pop_density_weights[[pop_w]][1]) + 
                 (prox_mm * pop_density_weights[[pop_w]][2]) + 
                 (pop_den_mm * pop_density_weights[[pop_w]][3]))
      
      # liveability score final weighting
      li_weight['li_mm'] <- minmax(li_weight, li_score)
      
      # rank and decile
      li_weight <- li_weight %>% ungroup() %>% mutate(li_rank = min_rank(li_mm),
                                      li_decile = ntile(li_mm, 10))
      
      # print current iteration
      print(paste('div_w: ', div_w, ' prox_w: ', prox_w, 'pop_w: ', pop_w))
      
      # create a final output table
      li_final <- select(li_weight, nb_clus, li_score, li_mm, li_rank, li_decile)
      
      # add variable fields
      li_final$diversity_weight <- div_w
      li_final$proximity_weight <- prox_w
      li_final$pop_density_weight <- pop_w
      li_final$normalisation <- 'minmax'
      
      ua_output_mm <- rbind(ua_output_mm, li_final)
      
      
    }
    
  }
}

```


### 6.2.3. Standardisation
```{r 72 scenarios - standardisation}
ua_output_ss <- tibble()

# run for loop - creates 72 versions of LI and store results for each neighbourhood

for (div_w in seq(1, length(diversity_weights))) {
  
  # diversity subindicators
  diversity <- df[1:6]
  
  # standardisation on each diversity subindicator
  diversity[,'d_c_ss'] <- ss(diversity, d_c)
  diversity[,'d_edu_ss'] <- ss(diversity, d_edu)
  diversity[,'d_enter_ss'] <- ss(diversity, d_enter)
  diversity[,'d_l_ss'] <- ss(diversity, d_l)
  diversity[,'d_h_ss'] <- ss(diversity, d_h)
  
  # diversity domain score
  div_weight <- diversity %>%
    mutate(div_score = (d_c_ss * diversity_weights[[div_w]][1]) + 
             (d_edu_ss * diversity_weights[[div_w]][2]) + 
             (d_enter_ss * diversity_weights[[div_w]][3]) + 
             (d_l_ss * diversity_weights[[div_w]][4]) + 
             (d_h_ss * diversity_weights[[div_w]][5]))
  
  # diversity final weighting
  div_weight['div_ss'] <- ss(div_weight, div_score)
  div_fin <- select(div_weight, nb_clus, div_ss)
  
  
  
  for (prox_w in seq(1, length(proximity_weights))) {
    
    # proximity subindicator
    proximity <- df[c(1, 7:11)]
    
    # standardisation on each proximity indicator
    proximity[,'p_c_ss'] <- ss_proximity(proximity, p_c)
    proximity[,'p_edu_ss'] <- ss_proximity(proximity, p_edu)
    proximity[,'p_enter_ss'] <- ss_proximity(proximity, p_enter)
    proximity[,'p_l_ss'] <- ss_proximity(proximity, p_l)
    proximity[,'p_h_ss'] <- ss_proximity(proximity, p_h)
    
    # proximity domain score
    prox_weight <- proximity %>%
      mutate(prox_score = (p_c_ss * proximity_weights[[prox_w]][1]) + 
               (p_edu_ss * proximity_weights[[prox_w]][2]) + 
               (p_enter_ss * proximity_weights[[prox_w]][3]) + 
               (p_l_ss * proximity_weights[[prox_w]][4]) + 
               (p_h_ss * proximity_weights[[prox_w]][5]))
    
    # proximity final weighting
    prox_weight['prox_ss'] <- ss(prox_weight, prox_score)
    prox_fin <- select(prox_weight, nb_clus, prox_ss)
    
    
    
    for (pop_w in seq(1, length(pop_density_weights))) {
      
      # population density
      pop_density <- df[c(1, 12)]
      
      # standardisation on population density
      pop_density['pop_den_ss'] <- ss(pop_density, pop_density)
      
      
      # combine three domains
      div_prox <- left_join(div_fin,
                            prox_fin,
                            by = 'nb_clus')
      
      div_prox_pop <- left_join(div_prox,
                                pop_density,
                                by = 'nb_clus')
      
      # final score
      li_weight <- div_prox_pop %>% 
        mutate(li_score = (div_ss * pop_density_weights[[pop_w]][1]) + 
                 (prox_ss * pop_density_weights[[pop_w]][2]) + 
                 (pop_den_ss * pop_density_weights[[pop_w]][3]))
      
      # liveability score final weighting
      li_weight['li_ss'] <- ss(li_weight, li_score)
      
      # rank and decile
      li_weight <- li_weight %>% ungroup() %>% mutate(li_rank = min_rank(li_ss),
                                      li_decile = ntile(li_ss, 10))
      
      # print current iteration
      print(paste('div_w: ', div_w, ' prox_w: ', prox_w, 'pop_w: ', pop_w))
      
      # create a final output table
      li_final <- select(li_weight, nb_clus, li_score, li_ss, li_rank, li_decile)
      
      # add variable fields
      li_final$diversity_weight <- div_w
      li_final$proximity_weight <- prox_w
      li_final$pop_density_weight <- pop_w
      li_final$normalisation <- 'ss'
      
      ua_output_ss <- rbind(ua_output_ss, li_final)
      
      
    }
    
  }
}

```


### 6.2.4. Robust normalisation
```{r 72 scenarios - robust normalisation}
ua_output_rs <- tibble()

# run for loop - creates 72 versions of LI and store results for each neighbourhood

for (div_w in seq(1, length(diversity_weights))) {
  
  # diversity subindicators
  diversity <- df[1:6]
  
  # robust normalisation on each diversity subindicator
  diversity[,'d_c_rs'] <- robust(diversity, d_c)
  diversity[,'d_edu_rs'] <- robust(diversity, d_edu)
  diversity[,'d_enter_rs'] <- robust(diversity, d_enter)
  diversity[,'d_l_rs'] <- robust(diversity, d_l)
  diversity[,'d_h_rs'] <- robust(diversity, d_h)
  
  # diversity domain score
  div_weight <- diversity %>%
    mutate(div_score = (d_c_rs * diversity_weights[[div_w]][1]) + 
             (d_edu_rs * diversity_weights[[div_w]][2]) + 
             (d_enter_rs * diversity_weights[[div_w]][3]) + 
             (d_l_rs * diversity_weights[[div_w]][4]) + 
             (d_h_rs * diversity_weights[[div_w]][5]))
  
  # diversity final weighting
  div_weight['div_rs'] <- robust(div_weight, div_score)
  div_fin <- select(div_weight, nb_clus, div_rs)
  
  
  
  for (prox_w in seq(1, length(proximity_weights))) {
    
    # proximity subindicator
    proximity <- df[c(1, 7:11)]
    
    # robust normalisation on each proximity indicator
    proximity[,'p_c_rs'] <- robust_proximity(proximity, p_c)
    proximity[,'p_edu_rs'] <- robust_proximity(proximity, p_edu)
    proximity[,'p_enter_rs'] <- robust_proximity(proximity, p_enter)
    proximity[,'p_l_rs'] <- robust_proximity(proximity, p_l)
    proximity[,'p_h_rs'] <- robust_proximity(proximity, p_h)
    
    # proximity domain score
    prox_weight <- proximity %>%
      mutate(prox_score = (p_c_rs * proximity_weights[[prox_w]][1]) + 
               (p_edu_rs * proximity_weights[[prox_w]][2]) + 
               (p_enter_rs * proximity_weights[[prox_w]][3]) + 
               (p_l_rs * proximity_weights[[prox_w]][4]) + 
               (p_h_rs * proximity_weights[[prox_w]][5]))
    
    # proximity final weighting
    prox_weight['prox_rs'] <- robust(prox_weight, prox_score)
    prox_fin <- select(prox_weight, nb_clus, prox_rs)
    
    
    
    for (pop_w in seq(1, length(pop_density_weights))) {
      
      # population density
      pop_density <- df[c(1, 12)]
      
      # standardisation on population density
      pop_density['pop_den_rs'] <- robust(pop_density, pop_density)
      
      
      # combine three domains
      div_prox <- left_join(div_fin,
                            prox_fin,
                            by = 'nb_clus')
      
      div_prox_pop <- left_join(div_prox,
                                pop_density,
                                by = 'nb_clus')
      
      # final score
      li_weight <- div_prox_pop %>% 
        mutate(li_score = (div_rs * pop_density_weights[[pop_w]][1]) + 
                 (prox_rs * pop_density_weights[[pop_w]][2]) + 
                 (pop_den_rs * pop_density_weights[[pop_w]][3]))
      
      # liveability score final weighting
      li_weight['li_rs'] <- robust(li_weight, li_score)
      
      # rank and decile
      li_weight <- li_weight %>% ungroup() %>% mutate(li_rank = min_rank(li_rs),
                                      li_decile = ntile(li_rs, 10))
      
      # print current iteration
      print(paste('div_w: ', div_w, ' prox_w: ', prox_w, 'pop_w: ', pop_w))
      
      # create a final output table
      li_final <- select(li_weight, nb_clus, li_score, li_rs, li_rank, li_decile)
      
      # add variable fields
      li_final$diversity_weight <- div_w
      li_final$proximity_weight <- prox_w
      li_final$pop_density_weight <- pop_w
      li_final$normalisation <- 'rs'
      
      ua_output_rs <- rbind(ua_output_rs, li_final)
      
      
    }
    
  }
}
```

### 6.2.5. Bind all scenarios
```{r}
# Assuming column names are consistent, adjust the data frames as needed
ua_output_exp <- rename(ua_output_exp, li_norm_score = li_exp)
ua_output_mm <- rename(ua_output_mm, li_norm_score = li_mm)
ua_output_rs <- rename(ua_output_rs, li_norm_score = li_rs)
ua_output_ss <- rename(ua_output_ss, li_norm_score = li_ss)

# Then combine them using rbind
ua_output <- rbind(ua_output_exp, ua_output_mm, ua_output_rs, ua_output_ss)

file_path = 'C:/Users/phily/liveability-index/data/uncertainty analysis/ua_data_final_450_1.csv'

write.csv(ua_output, file_path)

rm(ua_output_exp, ua_output_mm, ua_output_rs, ua_output_ss, div_w, pop_w, prox_w, exp_trans, minmax, minmax_proximity, rank_exp, rank_exp_proximity, robust, robust_proximity, ss, ss_proximity, norm_weights, proximity_weights, div_weight, div_fin, div_prox, div_prox_pop, diversity_weights, li_weight, pop_density, pop_density_weights, diversity, prox_fin, prox_weight, proximity)
gc()
```



## 6.3. Summary statistics for UA outputs
### 6.3.1. Reference ranking
```{r reference ranking}
# draw out baseline
baseline <- ua_output %>%
  filter(diversity_weight == 1, 
         proximity_weight == 1,
         pop_density_weight == 1,
         normalisation == 'rank_exp') %>%
  select(nb_clus, li_rank_b = li_rank)

# save reference ranking to csv for use in sensitivity analysis
write_csv(baseline, 'C:/Users/phily/liveability-index/data/uncertainty analysis/li_reference_rank_450.csv')
```

### 6.3.2. Summary stats
```{r}
# left join reference to full table, so we can work out each iterations score compared to the reference
ua_final <- ua_output %>%
  left_join(baseline, by = 'nb_clus') %>%
  mutate(li_rank_change = abs(li_rank - li_rank_b),
         rank_change_perc = li_rank_change / 399)

# save data
write_csv(ua_final, 'C:/Users/phily/liveability-index/data/uncertainty analysis/ua_all_tests_rank_change_450.csv')
```


### 6.3.3. Assign neighbourhoods to London Boroughs
The code below assigns each neighbourhood to boroughs it intersects with the most using spatial intersection. The neighbourhoods are constructed from the agglomeration of amenities, so they do not necessarily align with administrative boundary. Thus, some neighbourhoods overlap with multiple boroughs.

This procedure has been taken for the ease of interpretation and visualisation. This way, you can approximate the neighbourhoods' aggregation to boroughs.
```{r}
# read in london neighbourhood sf
nb_polygon

# London Boroughs
borough_sf <- st_read('C:/Users/phily/liveability-index/data/boundary/boros/London_Borough_Excluding_MHW.shp') %>% st_transform(27700)

# Step 1: Perform spatial intersection
intersection <- st_intersection(nb_polygon, borough_sf) # 719 observations

# Step 2: Determine the borough with the largest intersection area for each neighborhood
nb_polygon$LAD <- sapply(1:nrow(nb_polygon), function(i) {
  
  intersected_boroughs <- intersection$NAME[intersection$nb_clus == nb_polygon$nb_clus[i]]
  
  if (length(intersected_boroughs) > 0) {
    # Get the borough with the largest intersection area
    borough_with_max_area <- intersected_boroughs[which.max(intersection$geom[intersection$nb_clus == nb_polygon$nb_clus[i]] %>% st_area())]
    
    return(borough_with_max_area)
  } else {
    # Handle neighborhoods that do not intersect with any borough
    return(NA)
  }
})

# check whether there is any neighbourhood that is not assigned of boroughs
sum(is.na(nb_polygon$assigned_borough)) # all neighbourhoods have been assigned to boroughs

rm(intersection, borough_sf)
gc()
```

### 6.3.4. Data Join
```{r}
neighbourhood_lon <- nb_polygon %>% st_set_geometry(NULL)

# join to get LAD field and create unique test identifier to group by
results <- ua_output %>% 
  left_join(neighbourhood_lon, by = 'nb_clus') %>% 
  unite(test, c('normalisation', 'diversity_weight', 
                'proximity_weight', 'pop_density_weight'), 
        remove = FALSE)

# How many scenarios? 288 (6 * 6 * 2 * 4)
nrow(distinct(results, test))

```


## 6.5. Plots 
### 6.5.1. LAD Rankings for all simulations
```{r}
# Step 1: By grouping neighbourhoods by boroughs and test, we can calculate the aggregated average li_score for each borough
mean_scores_by_test <- results %>%
  group_by(LAD, test) %>%
  summarise(mean_li_score = mean(li_norm_score))

# Step 2: Calculate the rank of each borough under each of the 288 different test conditions. Boroughs with the highest li_score will have the rank of 33, while the one with the lowest will be assigned of a rank of 1.
lad_rank_for_288_tests <- mean_scores_by_test %>%
  group_by(test) %>%
  mutate(LAD_rank = min_rank(mean_li_score))
```


```{r overall LI Rank}
# Calculate the median rank for each borough
median_rank_by_borough <- lad_rank_for_288_tests %>%
  group_by(LAD) %>%
  summarize(median_rank = median(LAD_rank))

# Reorder the LAD factor levels based on the median rank
lad_rank_for_288_tests$LAD <- factor(
  lad_rank_for_288_tests$LAD,
  levels = median_rank_by_borough$LAD[order(median_rank_by_borough$median_rank)]
)

# Create the box plot for each borough with individual points
myplot1<- ggplot(lad_rank_for_288_tests, aes(x = LAD, y = LAD_rank)) +
  geom_boxplot(fill = "black", alpha = 0.1, width = 0.5) +
  geom_jitter(width = 0.2, size = 0.4, alpha = 0.3, color = "dodgerblue1") +
  labs(x = "Local Authority District (LAD)", y = "Rank") +
  ggtitle("Liveability Rank of Boroughs Under Different Simulations") +
  theme_bw() +  # Set the theme to a white background with black axes and grid lines
  theme(axis.text.x = element_text(angle = 50, hjust = 1, size = 11.5, family = 'sans'),
        axis.text.y = element_text(size = 11.5, family = 'sans'),
        axis.title = element_text(size = 14, family = 'sans'),
        plot.title = element_text(face = "bold", size = 20, hjust = 0.5 , family = 'sans', margin = margin(b = 20))) 
# Rotate x-axis labels for better readability

myplot1

ggsave('C:/Users/phily/liveability-index/output/uncertainty analysis/ua_rank_simulation_450_1.png', plot = myplot1, width = 12, height = 12, dpi = 300)

rm(lad_rank_for_288_tests, mean_scores_by_test, median_rank_by_borough, myplot1, results, ua_final, ua_output, neighbourhood_lon)
gc()
```



# 7. Results
## 7.1. Data Import
```{r}
# london boundary
london_boundary <- st_read('C:/Users/phily/liveability-index/data/boundary/london/london_boundary.shp') %>% st_transform(27700)

boros <- st_read('C:/Users/phily/liveability-index/data/boundary/boros/London_Borough_Excluding_MHW.shp') %>% st_transform(., 27700)

# LCI 
lci <- read_csv('C:/Users/phily/liveability-index/data/domain data/final_normalised_domain_indicator_nb_450.csv')

# neighbourhood polygon
nb_polygon <- st_read('C:/Users/phily/liveability-index/data/neighbourhood/final_neighbourhood_polygons.gpkg')

# assign LAD to nb_polygon
# Step 1: Perform spatial intersection
intersection <- st_intersection(nb_polygon, boros) # 719 observations

# Step 2: Determine the borough with the largest intersection area for each neighborhood
nb_polygon$LAD <- sapply(1:nrow(nb_polygon), function(i) {
  
  intersected_boroughs <- intersection$NAME[intersection$nb_clus == nb_polygon$nb_clus[i]]
  
  if (length(intersected_boroughs) > 0) {
    # Get the borough with the largest intersection area
    borough_with_max_area <- intersected_boroughs[which.max(intersection$geom[intersection$nb_clus == nb_polygon$nb_clus[i]] %>% st_area())]
    
    return(borough_with_max_area)
  } else {
    # Handle neighborhoods that do not intersect with any borough
    return(NA)
  }
})

neighbourhood_lon <- nb_polygon %>% st_set_geometry(NULL)

# left join nb_polygon and lci
lci_1 <- left_join(lci,
                   nb_polygon,
                   by = 'nb_clus') %>%
  st_as_sf()


gc()
```


```{r}
edges <- st_read("C:/Users/phily/Desktop/UCL/Term 2/UCL Dissertation/Dissertation_R/data/london_all.gpkg", layer = 'edges') %>%
  st_transform(., 27700)
edges <- edges[london_boundary,]
```


## 7.2. Liveability Index
```{r OVerall Liverability Index Plot}
# colour palette
library(RColorBrewer)
library(tmap)
library(tmaptools)

# Generate colours based on number of deciles and reverse them
# colours <- rev(brewer.pal(name = "Blues", n = nlevels(lci_1$li_decile)))

# Create the map
plot1 <- 
  tm_shape(edges) +
  tm_lines(col="black", alpha = 1, lwd = 0.3) +

  
  tm_shape(lci_1) +
  tm_fill(col = 'li_decile', 
              palette = 'Blues', 
              title = 'Decile',
              alpha = 0.6,
              style = 'cat') +
  tm_borders(lwd = 0.6, col = 'black') +
  tm_compass(type = "arrow", position = c('left', 'bottom'), size = 2) + 
  tm_scale_bar(text.size = 0.9, position = c("left", "bottom")) +
  tm_layout(legend.position = c("right", "bottom"), 
            #title= "Liveability Profiles of Greater London Neighbourhoods", 
            title.position = c('center', 'top'),
            title.size = 30,
            frame = FALSE,  # Remove the box surrounding the map
            inner.margins = 0.1)


# plot1
tmap_save(plot1, 'C:/Users/phily/liveability-index/output/results/overall_LI.png', height = 6, width = 8, dpi = 300)

rm(plot1)
gc()
```


## 7.3. Diversity Domain
```{r Diversity Plot}
# diversity 10 decile
lci_1 <- lci_1 %>% mutate(div_decile = ntile(diversity_score, 10))
# 
# # proximity 10 decile
lci_1 <- lci_1 %>% mutate(prox_decile = ntile(proximity_score, 10))
# 
# # population 10 decile
lci_1 <- lci_1 %>% mutate(pop_decile = ntile(pop_density_exp, 10))


# Create the map
plot2 <- 
  tm_shape(edges) +
  tm_lines(col="black", alpha = 1, lwd = 0.3) +
  
  tm_shape(lci_1) +
  tm_polygons(col = 'div_decile', 
              palette = "Blues",
              alpha = 0.6,
              title = 'Decile',
              style = 'cat') +
  tm_borders(lwd = 0.6, col = 'black') +
  tm_compass(type = "arrow", position = c('left', 'bottom'), size = 2) + 
  tm_scale_bar(text.size = 0.9, position = c("left", "bottom")) +
  tm_layout(legend.position = c("right", "bottom"), 
            # title= "Diversity Domain Profiles of Greater London Neighbourhoods", 
            title.position = c('center', 'top'),
            title.size = 30,
            frame = FALSE,  # Remove the box surrounding the map
            inner.margins = 0.1)


tmap_save(plot2, 'C:/Users/phily/liveability-index/output/results/diversity_domain.png', height = 6, width = 8, dpi = 300)

rm(plot2)
gc()
```
## 7.4. Proximity Domain
```{r Proximity Plot}
# Create the map
plot3 <- 
  tm_shape(edges) +
  tm_lines(col="black", alpha = 1, lwd = 0.3) +
  
  tm_shape(lci_1) +
  tm_polygons(col = 'prox_decile', 
              palette = "Blues",
              title = 'Decile',
              alpha = 0.6,
              style = 'cat') +
  tm_borders(lwd = 0.6, col = 'black') +
  tm_compass(type = "arrow", position = c('left', 'bottom'), size = 2) + 
  tm_scale_bar(text.size = 0.9, position = c("left", "bottom")) +
  tm_layout(legend.position = c("right", "bottom"), 
            #title= "Proximity Domain Profiles of 399 Neighbourhoods in London", 
            title.position = c('center', 'top'),
            title.size = 30,
            frame = FALSE,  # Remove the box surrounding the map
            inner.margins = 0.1)



tmap_save(plot3, 'C:/Users/phily/liveability-index/output/results/proximity_domain.png', height = 6, width = 8, dpi = 300)

rm(plot3)
gc()

```


## 7.5. Population Density Domain
```{r density Plot}
# Create the map
plot4 <- 
  tm_shape(edges) +
  tm_lines(col="black", alpha = 1, lwd = 0.3) +
  
  tm_shape(lci_1) +
  tm_polygons(col = 'pop_decile', 
              palette = "Blues",
              title = 'Decile',
              alpha = 0.6,
              style = 'cat') +
  tm_borders(lwd = 0.6, col = 'black') +
  tm_compass(type = "arrow", position = c('left', 'bottom'), size = 2) + 
  tm_scale_bar(text.size = 0.9, position = c("left", "bottom")) +
  tm_layout(legend.position = c("right", "bottom"), 
            #title= "Population Density Domain Profiles of 399 Neighbourhoods in London", 
            title.position = c('center', 'top'),
            title.size = 30,
            frame = FALSE,  # Remove the box surrounding the map
            inner.margins = 0.1)


#plot4
tmap_save(plot4, 'C:/Users/phily/liveability-index/output/results/density_domain.png', height = 6, width = 8, dpi = 300)

rm(plot4)
gc()
```

## 7.6. Decomposition of Indicators 
```{r}
# left join with LCI
lci_2 <- left_join(lci,
                   neighbourhood_lon,
                   by = 'nb_clus')

lci_per_borough <- lci_2 %>%
  group_by(LAD) %>%
  summarise(mean_d_c_score = mean(d_c_exp),
            mean_d_edu_score = mean(d_edu_exp),
            mean_d_enter_score = mean(d_enter_exp),
            mean_d_l_score = mean(d_l_exp),
            mean_d_h_score = mean(d_h_exp),
            mean_p_c_score = mean(p_c_exp),
            mean_p_edu_score = mean(p_edu_exp),
            mean_p_enter_score = mean(p_enter_exp),
            mean_p_l_score = mean(p_l_exp),
            mean_p_h_score = mean(p_h_exp),
            mean_den_score = mean(pop_density_exp),
            mean_li_score = mean(li_score)) %>% select(-mean_li_score)

# average liverability score sums upto individual indicator scores
# All indicator values are standardised and 1/11 is equal weights of individual indicators.

# multiply individual weights
lci_per_borough$mean_d_c_score <- lci_per_borough$mean_d_c_score * 1/11
lci_per_borough$mean_d_edu_score <- lci_per_borough$mean_d_edu_score * 1/11
lci_per_borough$mean_d_enter_score <- lci_per_borough$mean_d_enter_score * 1/11
lci_per_borough$mean_d_l_score <- lci_per_borough$mean_d_l_score * 1/11
lci_per_borough$mean_d_h_score <- lci_per_borough$mean_d_h_score * 1/11
lci_per_borough$mean_den_score <- lci_per_borough$mean_den_score * 1/11
lci_per_borough$mean_p_c_score <- lci_per_borough$mean_p_c_score * 1/11
lci_per_borough$mean_p_edu_score <- lci_per_borough$mean_p_edu_score * 1/11
lci_per_borough$mean_p_enter_score <- lci_per_borough$mean_p_enter_score * 1/11
lci_per_borough$mean_p_l_score <- lci_per_borough$mean_p_l_score * 1/11
lci_per_borough$mean_p_h_score <- lci_per_borough$mean_p_h_score * 1/11
```

### 7.7.1. LAD Decomposition Plot
```{r}
library(ggplot2)

# Define the legend names
legend_names <- c("mean_d_c_score" = "Diversity of Commerce",
                  "mean_d_edu_score" = "Diversity of Education",
                  "mean_d_enter_score" = "Diversity of Entertainment",
                  "mean_d_l_score" = "Diversity of Living",
                  "mean_d_h_score" = "Diversity of Healthcare",
                  "mean_p_c_score" = "Proximity to Commerce",
                  "mean_p_edu_score" = "Proximity to Education",
                  "mean_p_enter_score" = "Proximity to Entertainment",
                  "mean_p_l_score" = "Proximity to Living",
                  "mean_p_h_score" = "Proximity to Healthcare",
                  "mean_den_score" = "Population Density")

# Calculate the sum of indicators for each borough
lci_per_borough$sum_indicators <- rowSums(select(lci_per_borough, starts_with("mean_")))

# Reorder the levels of LAD based on the sum of indicators
lci_per_borough$LAD <- factor(lci_per_borough$LAD, levels = lci_per_borough$LAD[order(lci_per_borough$sum_indicators)])

# Reshape the data from wide to long format
lci_long <- lci_per_borough %>%
  pivot_longer(cols = starts_with("mean_"), names_to = "Indicator", values_to = "value")
```




```{r LAD Decomposed}
# Create the stacked bar chart
library(RColorBrewer)
# Define palettes (light → dark)
div_colors <- brewer.pal(6, "Greens")
prox_colors <- brewer.pal(5, "Blues")

# Assign manually so darkest = Living, lightest = Commerce
indicator_colors <- c(
  "mean_d_c_score"   = div_colors[6], # darkest
  "mean_d_edu_score" = div_colors[5],
  "mean_d_enter_score" = div_colors[4],
  "mean_d_h_score"   = div_colors[3],
  "mean_d_l_score"   = div_colors[2],  
  
  "mean_p_c_score"   = prox_colors[1],
  "mean_p_edu_score" = prox_colors[2],
  "mean_p_enter_score" = prox_colors[3],
  "mean_p_h_score"   = prox_colors[4],
  "mean_p_l_score"   = prox_colors[5],  # darkest
  
  "mean_den_score"   = "#FDB462"        # orange
)

# Plot
plot5 <- ggplot(lci_long, aes(x = LAD, y = value, fill = Indicator)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(x = "London Boroughs (LAD)", y = "Indicator Value") +
  ggtitle("Decomposition of Liveability Index into 11 Indicators") +
  scale_fill_manual(values = indicator_colors, labels = legend_names, name = "Indicator") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1, size = 11.5, family = 'sans'),
        axis.text.y = element_text(size = 11.5, family = 'sans'),
        axis.title = element_text(size = 14, family = 'sans'),
        plot.title = element_text(face = "bold", hjust = 0.5, size = 15, family = 'sans'),
        plot.margin = margin(10, 0, 10, 0),
        panel.background = element_rect(fill = "white", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA))

plot5

ggsave('C:/Users/phily/liveability-index/output/results/decomposed_LI.png', plot5, height = 10, width = 10, dpi = 300)

rm(plot5)
gc()
```

### 7.7.2 Normalisation of Indicator Scores - Top 3 Boroughs
```{r Top 3 Decomposed}
# Find the maximum score for each indicator
max_scores <- lci_long %>%
  group_by(Indicator) %>%
  summarise(max_score = max(value))

# Calculate the proportion of each indicator score relative to the maximum score
lci_long <- lci_long %>%
  left_join(max_scores, by = "Indicator") %>%
  mutate(proportion = value / max_score)

# Select the top 3 boroughs with the highest sum of indicator scores
top_3_boroughs <- lci_per_borough %>%
  arrange(desc(sum_indicators)) %>%
  head(3)


# Filter data for the bottom 3 boroughs
top_3_data <- lci_long %>%
  filter(LAD %in% top_3_boroughs$LAD)


library(dplyr)

# Recode the Indicator column with custom names
top_3_data <- top_3_data %>%
  mutate(Indicator = recode(Indicator,
    "mean_d_c_score" = "Diversity of Commerce",
    "mean_d_edu_score" = "Diversity of Education",
    "mean_d_enter_score" = "Diversity of Entertainment",
    "mean_d_h_score" = "Diversity of Healthcare",
    "mean_d_l_score" = "Diversity of Living",
    "mean_den_score" = "Population Density",
    "mean_p_c_score" = "Proximity to Commerce",
    "mean_p_edu_score" = "Proximity to Education",
    "mean_p_enter_score" = "Proximity to Entertainment",
    "mean_p_h_score" = "Proximity to Healthcare",
    "mean_p_l_score" = "Proximity to Living"
  ))

# Define colors matching your recoded Indicator names
indicator_colors <- c(
  "Diversity of Commerce"     = div_colors[6],
  "Diversity of Education"   = div_colors[5],
  "Diversity of Entertainment" = div_colors[4],
  "Diversity of Healthcare"  = div_colors[3],
  "Diversity of Living"      = div_colors[2],
  
  "Proximity to Commerce"    = prox_colors[1],
  "Proximity to Education"   = prox_colors[2],
  "Proximity to Entertainment" = prox_colors[3],
  "Proximity to Healthcare"  = prox_colors[4],
  "Proximity to Living"      = prox_colors[5],
  
  "Population Density"       = "#FDB462"
)

# Reorder the Indicator factor manually
top_3_data$Indicator <- factor(top_3_data$Indicator,
                                  levels = c(
                                    
                                    # Proximity indicators next (bottom)
                                    "Proximity to Living",
                                    "Proximity to Healthcare",
                                    "Proximity to Entertainment",
                                    "Proximity to Education",
                                    "Proximity to Commerce",
                                    
                                    # Population density in the middle
                                    "Population Density",
                                    
                                    # Diversity indicators first (top)
                                    "Diversity of Living",
                                    "Diversity of Healthcare",
                                    "Diversity of Entertainment",
                                    "Diversity of Education",
                                    "Diversity of Commerce"

                                    
                                    
                                    
                                    
                                  ))

# plot top 3 boroughs
plot_combined <- ggplot(top_3_data, aes(x = Indicator, y = proportion, fill = Indicator)) +
  geom_bar(stat = "identity") +
  xlab('Indicator') +
  ylab("Indicator Proportion") +
  ggtitle("Decomposed Liveability Index for Three Highest-Scoring Boroughs") +
  scale_fill_manual(values = indicator_colors) +  # Use custom colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 1, size = 14, family = 'sans', margin = margin(t = 10, r = 0, b = 10, l = 0)),
    axis.text.y = element_text(hjust = 1, size = 14, family = 'sans', margin = margin(t = 10, r = 0, b = 10, l = 0)),
    strip.background = element_rect(color = "black", fill = "grey90"),
    strip.text = element_text(size = 20, family = 'sans', face = 'bold'),
    axis.title = element_text(size = 20, family = 'sans', margin = margin(t = 10, r = 0, b = 10, l = 0), face = 'bold'),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 26, family = 'sans', margin = margin(b = 20)),
    plot.margin = unit(c(1, 1, 0.5, 1), "lines"),
    panel.spacing = unit(1.5, "lines")
  ) +
  facet_grid(. ~ LAD, scales = "free_y", space = "free_y", switch = "y") +
  scale_y_continuous(limits = c(0, 1)) +
  coord_flip() +
  guides(fill = FALSE)

# save
ggsave('C:/Users/phily/liveability-index/output/results/top3_decomposed_500.png' ,plot_combined, dpi = 300, width = 18, height = 10)
```

### 7.7.3. Normalisation of Indicator Scores - Bottom 3 Boroughs
```{r Bottom 3 Decomposed}
bottom_3_boroughs <- lci_per_borough %>%
  arrange(desc(sum_indicators)) %>%
  tail(3)


# Filter data for the bottom 3 boroughs
bottom_3_data <- lci_long %>%
  filter(LAD %in% bottom_3_boroughs$LAD)


library(dplyr)

# Recode the Indicator column with custom names
bottom_3_data <- bottom_3_data %>%
  mutate(Indicator = recode(Indicator,
    "mean_d_c_score" = "Diversity of Commerce",
    "mean_d_edu_score" = "Diversity of Education",
    "mean_d_enter_score" = "Diversity of Entertainment",
    "mean_d_h_score" = "Diversity of Healthcare",
    "mean_d_l_score" = "Diversity of Living",
    "mean_den_score" = "Population Density",
    "mean_p_c_score" = "Proximity to Commerce",
    "mean_p_edu_score" = "Proximity to Education",
    "mean_p_enter_score" = "Proximity to Entertainment",
    "mean_p_h_score" = "Proximity to Healthcare",
    "mean_p_l_score" = "Proximity to Living"
  ))

# Define colors matching your recoded Indicator names
indicator_colors <- c(
  "Diversity of Commerce"     = div_colors[6],
  "Diversity of Education"   = div_colors[5],
  "Diversity of Entertainment" = div_colors[4],
  "Diversity of Healthcare"  = div_colors[3],
  "Diversity of Living"      = div_colors[2],
  
  "Proximity to Commerce"    = prox_colors[1],
  "Proximity to Education"   = prox_colors[2],
  "Proximity to Entertainment" = prox_colors[3],
  "Proximity to Healthcare"  = prox_colors[4],
  "Proximity to Living"      = prox_colors[5],
  
  "Population Density"       = "#FDB462"
)

# Reorder the Indicator factor manually
bottom_3_data$Indicator <- factor(bottom_3_data$Indicator,
                                  levels = c(
                                    
                                    # Proximity indicators next (bottom)
                                    "Proximity to Living",
                                    "Proximity to Healthcare",
                                    "Proximity to Entertainment",
                                    "Proximity to Education",
                                    "Proximity to Commerce",
                                    
                                    # Population density in the middle
                                    "Population Density",
                                    
                                    # Diversity indicators first (top)
                                    "Diversity of Living",
                                    "Diversity of Healthcare",
                                    "Diversity of Entertainment",
                                    "Diversity of Education",
                                    "Diversity of Commerce"

                                    
                                    
                                    
                                    
                                  ))



# Create a single plot with shared y-axis in one row and three columns
plot_combined <- ggplot(bottom_3_data, aes(x = Indicator, y = proportion, fill = Indicator)) +
  geom_bar(stat = "identity") +
  xlab('Indicator') +
  ylab("Indicator Proportion") +
  ggtitle("Decomposed Liveability Index for Three Lowest-Scoring Boroughs") +
  scale_fill_manual(values = indicator_colors) +  # Use custom colors
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 1, size = 14, family = 'sans', margin = margin(t = 10, r = 0, b = 10, l = 0)),
    axis.text.y = element_text(hjust = 1, size = 14, family = 'sans', margin = margin(t = 10, r = 0, b = 10, l = 0)),
    strip.background = element_rect(color = "black", fill = "grey90"),
    strip.text = element_text(size = 20, family = 'sans', face = 'bold'),
    axis.title = element_text(size = 20, family = 'sans', margin = margin(t = 10, r = 0, b = 10, l = 0), face = 'bold'),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 26, family = 'sans', margin = margin(b = 20)),
    plot.margin = unit(c(1, 1, 0.5, 1), "lines"),
    panel.spacing = unit(1.5, "lines")
  ) +
  facet_grid(. ~ LAD, scales = "free_y", space = "free_y", switch = "y") +
  scale_y_continuous(limits = c(0, 1)) +
  coord_flip() +
  guides(fill = FALSE)

# Display the combined plot
print(plot_combined)


ggsave('C:/Users/phily/liveability-index/output/results/bottom3_decomposed_500.png' ,plot_combined, dpi = 300, width = 18, height = 10)

rm(plot_combined)
gc()
```

# 8. Sensitivity Analysis
## 8.1. Customised functions
```{r dirichlet functions}
# DIRICHLET WEIGHT SAMPLES
# move last 'n' items from a list to the front
lCycle <- function(x, n = 1) {
  if (n == 0) x else c(tail(x, -n), head(x, n))
}

# this function creates a table of nsamples from the dirichlet distribution for nweights (no. of indicators)
# each of the nweights will be treated with the alpha values in turn, while the others remain at 1
dirSampleSpread <- function(alpha_vals, nsamples, nweights){
  
  out <- data.frame()
  # loop through alpha values
  for (a in seq(1, length(alpha_vals))){
    
    # create a vector 'weights' where all weights are set to 1, except for the last one, which is set to the current alpha value
    weights <- c(rep(1, (nweights - 1)), alpha_vals[a])

    for (w in seq(0, nweights - 1)){
      
      # By cycling the weights, each indicator gets a turn to be weighted by alpha value, while other weights remain at 1
      w_list <- lCycle(weights, w)
      
      # generate dirichlet samples using the 'w_list' as the parameter
      weight_samples_m <- rdirichlet(nsamples, w_list)
      # turn to dataframe
      weight_samples <- as.data.frame(weight_samples_m)
      # rename weight variables and then store alpha value
      names(weight_samples) <- paste('w', seq(1, nweights), sep = '')
      
      weight_samples$alpha <-alpha_vals[a]
      
      # tracking focused weight - 'w_focus' indicates which weight in the original sequence was assigned the alpha value in the current iteration
      weight_samples$w_focus <- nweights - w
      
      # # bind to output table
      out <- rbind(out, weight_samples)
    }
  }
  out
}
```

```{r functions for diversity}
# exponential transformation function
# The higher the value is, the higher transformed value it gets.
# As higher value indicates the better diversity of neighbourhood
exp_trans <- function(var){
  -23 * log(1 - (var * (1 - (exp(-100/23)))))
}

# function to rank and exponentially transform a variable in a given table
rank_exp <- function(df, var){
  var <- enquo(var)

  df %>%
    ungroup() %>%   # <- critical
    mutate(rank   = min_rank(!! var),
           rank_s = scales::rescale(rank, to = c(0, 1)),
           exp    = exp_trans(rank_s)) %>%
    pull(exp)
}
```

```{r functions for proximity}
# exponential transformation function
# Unlike diversity, the shorter the distance is, the better the proximity is.
# So, we have to modify the previous function.
# We have to reverse the rank as this is opposite of diversity
# Exponential transformation function
exp_trans <- function(var){
  -23 * log(1 - (var * (1 - (exp(-100/23)))))
}

# Function to rank and exponentially transform a variable in a given table
rank_exp_proximity <- function(df, var){
  var <- enquo(var)
  
  df %>%
    ungroup() %>%  # remove any grouping
    mutate(
      rank = min_rank(!!var),
      reversed_rank = max(rank) - rank + 1,          # reverse rank for proximity
      rank_s = scales::rescale(reversed_rank, to = c(0, 1)), # rescale to 0-1
      exp = exp_trans(rank_s)
    ) %>%
    pull(exp)  # return as a numeric vector
}
```


## 8.2. Data
It is important to import all indicator dataset before normalisation and aggregation step because we are trying to assess how weighting changes in one indicator affects the whole liveability ranking.
```{r data import}
library(gtools)

# read in data
df <- read.csv('C:/Users/phily/liveability-index/data/domain data/raw_domain_indicator_nb_450.csv')
baseline <- read.csv('C:/Users/phily/liveability-index/data/uncertainty analysis/li_reference_rank_450.csv')

diversity <- df[1:6]
proximity <- df[c(1, 7:11)]
pop_density <- df[c(1, 12)]
```

### 8.2.1. Dirichlet Sampling for Proximity and Diversity domain
```{r sampling for proximity and diversity domains - 5 indicators}
library(gtools)

# create table of sample weights - 5 variables, 1000 samples for each iteration, alpha ranging from 1-10
weights <- dirSampleSpread(alpha_vals = seq(1, 10, 1), nsamples = 1000, nweights = 5)

# sense check output
weights <- arrange(weights, w_focus, alpha)
nrow(weights)
head(weights)
head(filter(weights, w_focus == 1))
head(filter(weights, w_focus == 2))
head(filter(weights, w_focus == 3))
head(filter(weights, w_focus == 4))
head(filter(weights, w_focus == 5))
weights %>% group_by(w_focus, alpha) %>% summarise(n = n())

```

#### 8.2.1.1. testing SA for Diversity domain
```{r proximity and population domain}
# exponential transformation on each proximity indicator
proximity[,'p_c_exp'] <- rank_exp_proximity(proximity, p_c)
proximity[,'p_enter_exp'] <- rank_exp_proximity(proximity, p_enter)
proximity[,'p_edu_exp'] <- rank_exp_proximity(proximity, p_edu)
proximity[,'p_l_exp'] <- rank_exp_proximity(proximity, p_l)
proximity[,'p_h_exp'] <- rank_exp_proximity(proximity, p_h)

# Proximity domain - final equal weighting
prox_weight <- proximity %>%
  mutate(proximity_score = (p_c_exp / 5) + (p_edu_exp / 5) +
           (p_enter_exp / 5) + (p_l_exp / 5) + (p_h_exp / 5))

prox_weight['proximity_exp'] <- rank_exp(prox_weight, proximity_score)
prox_fin <- select(prox_weight, nb_clus, proximity_exp)

# population density
pop_density['pop_density_exp'] <- rank_exp(pop_density, pop_density)
pop_fin <- pop_density %>% select(nb_clus, pop_density_exp)

# merge prox_fin and pop_fin
prox_pop <- prox_fin %>% left_join(pop_fin, by = 'nb_clus')
```


```{r diversity domain}
# exponential transformation on each diversity indicator
diversity[,'d_c_exp'] <- rank_exp(diversity, d_c)
diversity[,'d_edu_exp'] <- rank_exp(diversity, d_edu)
diversity[,'d_enter_exp'] <- rank_exp(diversity, d_enter)
diversity[,'d_l_exp'] <- rank_exp(diversity, d_l)
diversity[,'d_h_exp'] <- rank_exp(diversity, d_h)

# loop for sensitivity
score_list <- NULL

for (w in seq(1, nrow(weights))) {
  
  # use weights from list to calculate score
  diversity <- diversity %>%
    mutate(diversity_score = (d_c_exp * weights[w, 'w1']) + 
             (d_edu_exp * weights[w, 'w2']) + 
             (d_enter_exp * weights[w, 'w3']) + 
             (d_l_exp * weights[w, 'w4']) + 
             (d_h_exp * weights[w, 'w5']))
  
  # transform to exponential and create final table
  diversity['diversity_exp'] <- rank_exp(diversity, diversity_score)
  div_fin <- select(diversity, nb_clus, diversity_exp)
  
  # final index
  li_temp <- prox_pop %>% 
    left_join(div_fin, by = 'nb_clus')
  
  li_w1 <- li_temp %>%
    mutate(li_score = (proximity_exp / 3) + (pop_density_exp / 3) + 
             (diversity_exp / 3))
  
  # calculate the total score
  li_w1['li_exp'] <- rank_exp(li_w1, li_score)

  # calculate ranking
  li_w1 <- li_w1 %>% ungroup() %>% mutate(li_rank = min_rank(li_exp))
  
  # create final output table - joining to baseline to calculate rank change per neighbourhood
  li_fin <- select(li_w1, nb_clus, li_rank) %>%
    left_join(baseline, by = 'nb_clus') %>%
    mutate(rank_change = abs(li_rank - li_rank_b))
  
  mean_r_change <- mean(li_fin$rank_change)
  score_list <- append(score_list, mean_r_change)
  
}

# append score list of mean rank change per test back to weights table
weights$rank_change <- score_list
weights$rank_change_percent <- score_list / 399

file_path = 'C:/Users/phily/liveability-index/data/sensitivity analysis/sa_dirichlet_diversity_results_1000_1.csv'

write.csv(weights, file_path, row.names = FALSE)
```



#### 8.2.1.2. testing SA for Proximity domain
```{r diversity and population domain}
# exponential transformation on each diversity indicator
diversity[,'d_c_exp'] <- rank_exp(diversity, d_c)
diversity[,'d_edu_exp'] <- rank_exp(diversity, d_edu)
diversity[,'d_enter_exp'] <- rank_exp(diversity, d_enter)
diversity[,'d_l_exp'] <- rank_exp(diversity, d_l)
diversity[,'d_h_exp'] <- rank_exp(diversity, d_h)

# diversity domain - final equal weighting
div_weight <- diversity %>%
  mutate(diversity_score = (d_c_exp / 5) + (d_edu_exp / 5) +
           (d_enter_exp / 5) + (d_l_exp / 5) + (d_h_exp / 5))

div_weight['diversity_exp'] <- rank_exp(div_weight, diversity_score)
div_fin <- select(div_weight, nb_clus, diversity_exp)

# population density
pop_density['pop_density_exp'] <- rank_exp(pop_density, pop_density)
pop_fin <- pop_density %>% select(nb_clus, pop_density_exp)

# merge prox_fin and pop_fin
div_pop <- div_fin %>% left_join(pop_fin, by = 'nb_clus')
```


```{r proximity domain}
# exponential transformation on each proximity indicator
proximity[,'p_c_exp'] <- rank_exp_proximity(proximity, p_c)
proximity[,'p_edu_exp'] <- rank_exp_proximity(proximity, p_edu)
proximity[,'p_enter_exp'] <- rank_exp_proximity(proximity, p_enter)
proximity[,'p_l_exp'] <- rank_exp_proximity(proximity, p_l)
proximity[,'p_h_exp'] <- rank_exp_proximity(proximity, p_h)

# loop for sensitivity
score_list <- NULL

for (w in seq(1, nrow(weights))) {
  
  # use weights from list to calculate score
  proximity <- proximity %>%
    mutate(proximity_score = (p_c_exp * weights[w, 'w1']) + 
             (p_edu_exp * weights[w, 'w2']) + 
             (p_enter_exp * weights[w, 'w3']) + 
             (p_l_exp * weights[w, 'w4']) + 
             (p_h_exp * weights[w, 'w5']))
  
  # transform to exponential and create final table
  proximity['proximity_exp'] <- rank_exp(proximity, proximity_score)
  prox_fin <- select(proximity, nb_clus, proximity_exp)
  
  # final index
  li_temp <- div_pop %>% 
    left_join(prox_fin, by = 'nb_clus')
  
  li_w1 <- li_temp %>%
    mutate(li_score = (proximity_exp / 3) + (pop_density_exp / 3) + 
             (diversity_exp / 3))
  
  # calculate the total score
  li_w1['li_exp'] <- rank_exp(li_w1, li_score)

  # calculate ranking
  li_w1 <- li_w1 %>% ungroup() %>% mutate(li_rank = min_rank(li_exp))
  
  
  # create final output table - joining to baseline to calculate rank change per neighbourhood
  li_fin <- select(li_w1, nb_clus, li_rank) %>%
    left_join(baseline, by = 'nb_clus') %>%
    mutate(rank_change = abs(li_rank - li_rank_b))
  
  # calculate the mean rank_change in li_fin data from each sample simulation
  mean_r_change <- mean(li_fin$rank_change)
  
  # store it in the list
  score_list <- append(score_list, mean_r_change)
  
}

# append score list of mean rank change per test back to weights table
weights$rank_change <- score_list
weights$rank_change_percent <- score_list / 399

file_path = 'C:/Users/phily/liveability-index/data/sensitivity analysis/sa_dirichlet_proximity_results_1000_1.csv'

write.csv(weights, file_path, row.names = FALSE)
```



### 8.2.2. Dirichlet Sampling for LI domains
```{r sampling for proximity and diversity domains - 3 domains}
# create table of sample weights - 3 variables, 1000 samples for each iteration, alpha ranging from 1-10
weights <- dirSampleSpread(alpha_vals = seq(1, 10, 1), nsamples = 1000, nweights = 3)

# sense check output
weights <- arrange(weights, w_focus, alpha)
nrow(weights)
head(weights)
head(filter(weights, w_focus == 1))
head(filter(weights, w_focus == 2))
head(filter(weights, w_focus == 3))
weights %>% group_by(w_focus, alpha) %>% summarise(n = n())
```

```{r all domains}
# exponential transformation on each diversity indicator
diversity[,'d_c_exp'] <- rank_exp(diversity, d_c)
diversity[,'d_edu_exp'] <- rank_exp(diversity, d_edu)
diversity[,'d_enter_exp'] <- rank_exp(diversity, d_enter)
diversity[,'d_l_exp'] <- rank_exp(diversity, d_l)
diversity[,'d_h_exp'] <- rank_exp(diversity, d_h)

# diversity domain - final equal weighting
div_weight <- diversity %>%
  mutate(diversity_score = (d_c_exp / 5) + (d_edu_exp / 5) +
           (d_enter_exp / 5) + (d_l_exp / 5) + (d_h_exp / 5))

div_weight['diversity_exp'] <- rank_exp(div_weight, diversity_score)
div_fin <- select(div_weight, nb_clus, diversity_exp)



# exponential transformation on each proximity indicator
proximity[,'p_c_exp'] <- rank_exp_proximity(proximity, p_c)
proximity[,'p_enter_exp'] <- rank_exp_proximity(proximity, p_enter)
proximity[,'p_edu_exp'] <- rank_exp_proximity(proximity, p_edu)
proximity[,'p_l_exp'] <- rank_exp_proximity(proximity, p_l)
proximity[,'p_h_exp'] <- rank_exp_proximity(proximity, p_h)

# Proximity domain - final equal weighting
prox_weight <- proximity %>%
  mutate(proximity_score = (p_c_exp / 5) + (p_edu_exp / 5) +
           (p_enter_exp / 5) + (p_l_exp / 5) + (p_h_exp / 5))

prox_weight['proximity_exp'] <- rank_exp(prox_weight, proximity_score)
prox_fin <- select(prox_weight, nb_clus, proximity_exp)

# population density
pop_density['pop_density_exp'] <- rank_exp(pop_density, pop_density)
pop_fin <- pop_density %>% select(nb_clus, pop_density_exp)


# all domains
all_fin <- div_fin %>% left_join(prox_fin, by = 'nb_clus') %>% 
  left_join(pop_fin, by = 'nb_clus')
  
```

```{r loop}
# loop for sensitivity
score_list <- NULL

for (w in seq(1, nrow(weights))) {
  
  # use weights from list to calculate score
  all_fin <- all_fin %>%
    mutate(li_score = (diversity_exp * weights[w, 'w1']) + 
             (proximity_exp * weights[w, 'w2']) + 
             (pop_density_exp * weights[w, 'w3']))
  
  # transform to exponential and create final table
  all_fin['li_exp'] <- rank_exp(all_fin, li_score)
  
  # final index
  li_w1 <- all_fin %>% ungroup() %>% mutate(li_rank = min_rank(li_exp))
  
  # create final output table - joining to baseline to calculate rank change per neighbourhood
  li_fin <- select(li_w1, nb_clus, li_rank) %>%
    left_join(baseline, by = 'nb_clus') %>%
    mutate(rank_change = abs(li_rank - li_rank_b))
  
  mean_r_change <- mean(li_fin$rank_change)
  score_list <- append(score_list, mean_r_change)
  
}

# append score list of mean rank change per test back to weights table
weights$rank_change <- score_list
weights$rank_change_percent <- score_list / 399

file_path = 'C:/Users/phily/liveability-index/data/sensitivity analysis/sa_dirichlet_results_1000_1.csv'

write.csv(weights, file_path, row.names = FALSE)
```

## 8.3. Visualisation
### 8.3.1. All Domains
```{r data import}
# data read
data <- read.csv('C:/Users/phily/liveability-index/data/sensitivity analysis/sa_dirichlet_results_1000_1.csv')

# reshape data - grouped by alpha and w_focus so we can calculate mean rank change for each domain
data_summary <- data %>%
  group_by(alpha, w_focus) %>%
  summarise(mean_rank_change = mean(rank_change_percent))

# Convert w_focus to factor and rename
data_summary$w_focus <- factor(data_summary$w_focus, levels = c(1, 2, 3), labels = c("Diversity", "Proximity", "Population Density"))

# Filter the data to include only rows where alpha increases by 1
markers <- data_summary %>%
  filter(alpha %% 1 == 0)

# Create the line graph with markers
plot_a <- ggplot(data_summary, aes(x = alpha, y = mean_rank_change, color = w_focus)) +
  geom_line() +
  geom_point(data = markers, aes(x = alpha, y = mean_rank_change), shape = 19, size = 3) +
  labs(#title = "Sensitivity Analysis - Domain Weights",
       x = "Alpha Value",
       y = "Mean Rank Change",
       color = "Liveability Domains") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(face = "bold"),
    legend.position = c(0.16, 0.83),  # (x,y) coordinates inside plot
    legend.background = element_rect(fill = "white", color = "black") # optional for readability
  )


# Display the plot
print(plot_a)

# Save the plot
ggsave(filename = "C:/Users/phily/liveability-index/output/sensitivity analysis/sensitivity_analysis_plot_1.png", plot = plot_a, width = 10, height = 6, dpi = 800)

```

### 8.3.2. Diversity Domains
```{r data import}
# data read
data_1 <- read.csv('C:/Users/phily/liveability-index/data/sensitivity analysis/sa_dirichlet_diversity_results_1000_1.csv')

# reshape data - grouped by alpha and w_focus so we can calculate mean rank change for each domain
data_summary_1 <- data_1 %>%
  group_by(alpha, w_focus) %>%
  summarise(mean_rank_change = mean(rank_change_percent))

# Convert w_focus to factor and rename
data_summary_1$w_focus <- factor(data_summary_1$w_focus, levels = c(1, 2, 3, 4, 5), labels = c("Diversity of Commerce", "Diversity of Education", "Diversity of Entertainment", 'Diversity of Living', 'Diversity of Healthcare'))

# Filter the data to include only rows where alpha increases by 1
markers <- data_summary_1 %>%
  filter(alpha %% 1 == 0)

# Create the line graph with markers
plot_b <- ggplot(data_summary_1, aes(x = alpha, y = mean_rank_change, color = w_focus)) +
  geom_line() +
  geom_point(data = markers, aes(x = alpha, y = mean_rank_change), shape = 19, size = 3) +  # Add markers
  labs(#title = "Sensitivity Analysis - Diversity Indicator Weights",
       x = "Alpha Value",
       y = "Mean Rank Change",
       color = "Diversity Indicator") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(face = "bold"),
    legend.position = c(0.18, 0.79),  # (x,y) coordinates inside plot
    legend.background = element_rect(fill = "white", color = "black") # optional for readability
  )

# Display the plot
print(plot_b)

# Save the plot
ggsave(filename = "C:/Users/phily/liveability-index/output/sensitivity analysis/sensitivity_analysis_diversity_plot_1.png", plot = plot_b, width = 10, height = 6, dpi = 800)
```

### 8.3.3. Proximity Domain Indicators
```{r data import}
# data read
data_2 <- read.csv('C:/Users/phily/liveability-index/data/sensitivity analysis/sa_dirichlet_proximity_results_1000_1.csv')

# reshape data - grouped by alpha and w_focus so we can calculate mean rank change for each domain
data_summary_2 <- data_2 %>%
  group_by(alpha, w_focus) %>%
  summarise(mean_rank_change = mean(rank_change_percent))

# Convert w_focus to factor and rename
data_summary_2$w_focus <- factor(data_summary_2$w_focus, levels = c(1, 2, 3, 4, 5), labels = c("Proximity to Commerce", "Proximity to Education", "Proximity to Entertainment", 'Proximity to Living', 'Proximity to Healthcare'))

# Filter the data to include only rows where alpha increases by 1
markers <- data_summary_2 %>%
  filter(alpha %% 1 == 0)


plot_c <- ggplot(data_summary_2, aes(x = alpha, y = mean_rank_change, color = w_focus)) +
  geom_line() +
  geom_point(data = markers, aes(x = alpha, y = mean_rank_change), shape = 19, size = 3) +
  scale_y_continuous(
    breaks = c(seq(0, max(data_summary_2$mean_rank_change), by = 0.01), 0.06),
    limits = c(0.03, 0.06)  # set y axis limits
  ) +
  labs(#title = "Sensitivity Analysis - Proximity Indicator Weights",
       x = "Alpha Value",
       y = "Mean Rank Change",
       color = "Proximity Indicator") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.title = element_text(face = "bold"),
    legend.position = c(0.18, 0.79),  # (x,y) coordinates inside plot
    legend.background = element_rect(fill = "white", color = "black") # optional for readability
  )



# Display the plot
print(plot_c)

# Save the plot
ggsave(filename = "C:/Users/phily/liveability-index/output/sensitivity analysis/sensitivity_analysis_proximity_plot_1.png", plot = plot_c, width = 10, height = 6, dpi = 800)
```





# 9. Validation against IMD
## 9.1. Data Import
```{r}
nb_polygon <- st_read('C:/Users/phily/liveability-index/data/neighbourhood/final_neighbourhood_polygons.gpkg')

hexagon_scores <- st_read('C:/Users/phily/liveability-index/data/IMD/hexagon_imd.geojson')

gc()
```
## 9.2. Projecting hexagon-level IMD score to neighbourhood-level
```{r GBS scores at neighbourhood level}
# calculate the area size of each neighbourhood
hexagon_scores$hex_area <- st_area(hexagon_scores)

# Perform spatial intersection
intersection <- st_intersection(hexagon_scores, nb_polygon)

# Calculate the proportion of each hexagon's area within each neighborhood polygon
intersection$area_proportion <- st_area(intersection) / intersection$hex_area

max(intersection$area_proportion)
min(intersection$area_proportion)

# Calculate the estimated population within each neighborhood polygon
intersection$estimated_imd_score <- intersection$weighted_avg_imd * intersection$area_proportion

# When sum of weights does not sum up to 1, weighted average is calculated by 
# Sum of each number multiplied by its weight / sum of all weights = weighted average
# Group by hexagon index
neighbourhood_imd <- intersection %>%
  group_by(nb_clus) %>%
  summarise(weighted_avg_imd = sum(estimated_imd_score) / sum(area_proportion))

min(neighbourhood_imd$weighted_avg_imd)
max(neighbourhood_imd$weighted_avg_imd)

# # save data - just those with populations
file_path = 'C:/Users/phily/liveability-index/data/IMD/neighbourhood_450_imd.geojson'
# #
st_write(neighbourhood_imd, file_path, driver = 'GeoJSON')
```

## 9.3. Correlation Plots
```{r read in data}
lci <- read_csv('C:/Users/phily/liveability-index/data/domain data/final_normalised_domain_indicator_nb_450.csv')

neighbourhood_imd <- st_read(file_path)
gc()
```

```{r merge data}
# Merge the dataframes based on the 'nb_clus' column
merged_data <- merge(neighbourhood_imd, lci, by = "nb_clus")

# drop units
library(units)

merged_data$weighted_avg_imd <- drop_units(merged_data$weighted_avg_imd)
```

```{r}
# correlation test
cor.test(merged_data$weighted_avg_imd, merged_data$li_exp,
         conf.level = 0.99)
```
The correlation between two indices are significantly different from 0 at the 1% level.

```{r linear regression}
# Calculate the R-squared and RMSE (Root Mean Squared Error)
lm_model <- lm(weighted_avg_imd ~ li_exp, data = merged_data)

# Calculate Adjusted R-squared - 0.56
summary(lm_model)$adj.r.squared
lm_model

# RMSE
predicted_values <- predict(lm_model)
rmse <- sqrt(mean((merged_data$li_exp - predicted_values)^2))
```
An increase of the li_exp by 1 unit decreases the IMD score by 0.018. The adjusted R-square of 0.56 between the Liveability Composite Index (LCI) and the IMD score. RMSE is 30.07.

```{r plot}
# Create a scatter plot
plot <- ggplot(merged_data, aes(x = li_exp, y = weighted_avg_imd)) +
  geom_jitter(alpha = 0.5) +
  geom_smooth(method = "lm", color = "dodgerblue2") +
  xlab("LI score") +
  ylab("Geographical Barriers Sub-domain score") +
  ggtitle("Correlation between LI score and the GBS score") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

file_path = 'C:/Users/phily/liveability-index/output/IMD/correlation_plot_450.png'
ggsave(file_path, plot, dpi = 800, width = 6, height = 6)
```

```{r}
library(ggplot2)

# Get the list of residuals
res <- resid(lm_model)

# Create a data frame for the plot
residuals_df <- data.frame(fitted = fitted(lm_model), residual = res)

# residuals_df$fitted <- drop_units(residuals_df$fitted)
# residuals_df$residual <- drop_units(residuals_df$residual)

# Create the plot using ggplot2
plot1 <- ggplot(residuals_df, aes(x = fitted, y = residual)) +
  geom_jitter(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residuals vs. Fitted Plot") +
  theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5)) 

file_path = 'C:/Users/phily/liveability-index/output/IMD/residual_plot_450.png'

ggsave(file_path, plot1, dpi = 300, width = 6, height = 6)
```

# 10. Active Travel
```{r}
library(sf)
library(dplyr)
# library(rgdal)
library(sp)
library(tidyverse)
```

## 10.1.  Average Footfall data at Uber H3 level 10
```{r footfall per hexagon}
# average footfall score per hexagon data
result <- read.csv("C:/Users/phily/liveability-index/data/footfall/average_footfall_per_hexagon.csv")

# read in UBER H3 level 10 hexagon cell data that covers london
london_hex <-st_read("C:/Users/phily/liveability-index/data/footfall/Greater_London.json") %>% 
  st_transform(27700)
```



```{r embed avg footfall data into London hexagon data}
# we need geometry column so we need to merge the result avg footfall data with hexagon grid sf data
hex_footfall <- left_join(london_hex,
                          result,
                          by = c('X_index' = 'hex_code')) %>% select(-c(2,3))


# replace na values with 0 as these na values indicate that people did not visit these hexagons.
hex_footfall['average_footfall'][is.na(hex_footfall['average_footfall'])] <- 0
sum(is.na(hex_footfall))

# max/min
max(hex_footfall$average_footfall) # 57.64
min(hex_footfall$average_footfall) # 0 originated from hexagons that people did not visit

# plot distribution of footfall data
ggplot(hex_footfall, aes(average_footfall)) +
  geom_histogram(binwidth = 0.5) +
  xlab('Average Normalised Footfall Score') +
  ylab('Frequency') +
  ggtitle('The Distribution of Average Normalised Footfall Score')
```


## 10.2. Footfall Score Projection to a Neighbourhood-level
```{r neighbourhood polygons}
nb_polygon <- st_read('C:/Users/phily/liveability-index/data/neighbourhood/final_neighbourhood_polygons.gpkg') %>% st_transform(27700)
```


```{r footfall scores at neighbourhood level}
# calculate the area size of each neighbourhood
hex_footfall$hex_area <- st_area(hex_footfall)

# Perform spatial intersection - 139011
intersection <- st_intersection(hex_footfall, nb_polygon)

# Calculate the proportion of each hexagon's area within each neighborhood polygon
intersection$area_proportion <- st_area(intersection) / intersection$hex_area

max(intersection$area_proportion)
min(intersection$area_proportion)

# Calculate the estimated footfall score within each neighborhood polygon
intersection$estimated_footfall_score <- intersection$average_footfall * intersection$area_proportion

# When sum of weights does not sum up to 1, weighted average is calculated by 
# Sum of each number multiplied by its weight / sum of all weights = weighted average
# Group by hexagon index
neighbourhood_fs <- intersection %>%
  group_by(nb_clus) %>%
  summarise(weighted_avg_footfall = sum(estimated_footfall_score) / sum(area_proportion))

min(neighbourhood_fs$weighted_avg_footfall) # 0.03
max(neighbourhood_fs$weighted_avg_footfall) # 1.45

# save data - just those with footfall score
file_path = 'C:/Users/phily/liveability-index/data/footfall/footfall_nb_450.geojson'

st_write(neighbourhood_fs, file_path, driver = 'GeoJSON')
gc()
```



## 10.3 Spatial Regression Models 
### 10.3.1. Linear Regression Model
```{r}
library(sf)
library(dplyr)
library(sp)
library(tidyverse)

# liveability index
lci <- read_csv('C:/Users/phily/liveability-index/data/domain data/final_normalised_domain_indicator_nb_450.csv')

neighbourhood_fs <- st_read('C:/Users/phily/liveability-index/data/footfall/footfall_nb_450.geojson')


# neighbourhood-level footfall score
neighbourhood_fs <- neighbourhood_fs %>% st_transform(27700)

# merge lci and footfall data
# Merge the dataframes based on the 'nb_clus' column
merged_data <- merge(neighbourhood_fs, lci, by = "nb_clus")

rm(lci, neighbourhood_fs)
gc()

# Select variables
myvars <- merged_data %>%
  select(weighted_avg_footfall, li_exp)


# check correlation
library('AICcmodavg')

# run OLS regression
model_ols <- lm(weighted_avg_footfall ~ li_exp,
                data = myvars)

# model summary
summary(model_ols)


# AIC and AICc - performance metrics
aic_ols <- AIC(model_ols)
n <- nrow(myvars)
k <- length(coef(model_ols))  # number of parameters incl. intercept

aicc_ols <- aic_ols + (2 * k * (k + 1)) / (n - k - 1)
```

```{r example}
summary(merged_data$weighted_avg_footfall)
summary(merged_data$li_exp)

# Liveability: 75th percentile (31.01), 25th percentile (6.52)
# footfall: median 0.487, mean 0.496
# OLS coef 0.0084

# difference in liveability
li_diff <- quantile(merged_data$li_exp, 0.75) - quantile(merged_data$li_exp, 0.25) 

# expected increase in footfall score
exp_inc_foot <- model_ols$coefficients["li_exp"] * li_diff

# 41.5%
relative_increase <- exp_inc_foot/mean(merged_data$weighted_avg_footfall)
```



### 10.3.2. Spatial Autocorrelation
```{r residual distribution}
# save residuals of OLS
merged_data <- merged_data %>%
  mutate(model_ols_res = residuals(model_ols))

# dataframe to sf
merged_data <- merged_data %>%
  st_as_sf()

# qtm(merged_data, fill = 'model_ols_res')

merged_data$model_ols_res_cat <- cut(
  merged_data$model_ols_res,
  breaks = c(-Inf, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8),
  labels = c("-0.6 to -0.4", "-0.4 to -0.2", "-0.2 to 0.0",
             "0.0 to 0.2", "0.2 to 0.4", "0.4 to 0.6", "0.6 to 0.8"),
  include.lowest = TRUE,
  right = FALSE
)


# Create the map
plot <- 

  tm_shape(merged_data) +
  tm_fill(col = 'model_ols_res_cat', 
              palette = "RdYlGn", 
              title = 'OLS Residuals',
              alpha = 0.6,
              style = 'cat') +
  tm_borders(lwd = 0.6, col = 'black') +
  tm_compass(type = "arrow", position = c(0.11, 0.1), size = 2) + 
  tm_scale_bar(text.size = 0.7, position = c(0,0.01)) +
  tm_layout(legend.position = c(0.88, 0.03), 
            title.position = c('center', 'top'),
            title.size = 30,
            frame = FALSE,  # Remove the box surrounding the map
            inner.margins = 0.1)


# Spatial autocorrelation plot
#tmap_save(plot, 'C:/Users/phily/liveability-index/output/results/spatial_autocorrelation.png', height = 6, width = 8, dpi = 300)


# calculate the centroids of all neighbourhoods in London
coordsW <- merged_data %>%
  st_centroid()%>%
  st_geometry()

# Now we need to generate a spatial weights matrix 
# We'll start with a simple binary matrix of queen's case neighbours
library(spdep)

# LNeigh_queen <- merged_data %>%
#   poly2nb(., queen=T)

#or nearest neighbours
knn_neigh <-coordsW %>%
  knearneigh(., k=4)

LNeigh_knn <- knn_neigh %>%
  knn2nb()

# plot them
# plot(LNeigh_queen, st_geometry(coordsW), col="red")
plot(LNeigh_knn, st_geometry(coordsW), col="blue")

#create a spatial weights list object from these weights
LNeigh.knn_4_weight <- LNeigh_knn %>%
  nb2listw(., style="W")

# Queen's case throws an error as one of neighbourhoods does not share any boundary with others
# LNeigh.queens_weight <- LNeigh_queen %>%
#   nb2listw(., style="W", zero.policy = T)

Nearest_neighbour <- merged_data %>%
  st_drop_geometry()%>%
  dplyr::select(model_ols_res)%>%
  pull()%>%
  moran.test(., LNeigh.knn_4_weight)

# Moran's I (-1 ~ +1, 0 being no spatial autocorrelation) - 0.3557 Moderate spatial autocorrelation in residuals
Nearest_neighbour
```


### 10.3.3. GWR
global regression model might not capture the relationship between variables that might vary locally - non-stationarity.
```{r}
library(spgwr)

coordsW2 <- st_coordinates(coordsW)

merged_data1 <- cbind(merged_data,coordsW2)

GWRbandwidth <- gwr.sel(weighted_avg_footfall ~ li_exp, 
                  data = merged_data1, 
                        coords=cbind(merged_data1$X, merged_data1$Y),
                  adapt=T)

GWRbandwidth

#run the gwr model
gwr.model = gwr(weighted_avg_footfall ~ li_exp,
                data = merged_data1, 
                coords=cbind(merged_data1$X, merged_data1$Y), 
                adapt=GWRbandwidth,
                #matrix output
                hatmatrix=TRUE,
                #standard error
                se.fit=TRUE)

# bandwidth 0.009 meaning 0.9% of all total spatial units should be used for the local regression based on the k-nearest neighbours, which equates to 3 neighbourhoods but too small sample size



#print the results of the model
gwr.model
```

```{r adjusted r-squared}
# Extract model components
RSS <- gwr.model$results$rss
n <- gwr.model$results$n
tr_S <- gwr.model$results$v1

# Compute TSS manually
y <- merged_data1$weighted_avg_footfall
TSS <- sum((y - mean(y))^2)

# Compute Adjusted R^2
adj_r2 <- 1 - ((RSS / (n - tr_S)) / (TSS / (n - 1)))
adj_r2
```



```{r gwr coefficients}
# Check coefficients
results <- as.data.frame(gwr.model$SDF)

#attach coefficients to original SF
merged_data2 <- merged_data %>%
  mutate(coef_li_exp = results$li_exp)

hist(merged_data2$coef_li_exp)

library(classInt)
set.seed(42)

test <- merged_data2 %>% select(coef_li_exp) %>%
  filter(coef_li_exp > 0)

# Quantile classification
quantile_breaks <- classIntervals(test$coef_li_exp, n = 4, style = "quantile")$brks

# manual classification based on quantile and histogram to separate negative coefs and for visualisation purpose
quantile_breaks <- c(-0.0006, 0, 0.0072, 0.0104, 0.0151, 0.025)

# Assign category labels
merged_data2$coef_li_exp_cat <- cut(merged_data2$coef_li_exp,
                                   breaks = quantile_breaks,
                                   include.lowest = TRUE,
                                   right = FALSE,
                                   labels = c("Below 0",
                                              "0 - 0.0072",
                                              "0.0072 - 0.0104",
                                              "0.0104 - 0.0151",
                                              "0.0151 - 0.025"))


# Quantile colour palette 
color_palette <- c("Below 0" = '#2166AC', 
                "0 - 0.0072" = '#67A9CF',
                "0.0072 - 0.0104" = '#B2E2B4',
                "0.0104 - 0.0151" = '#CCEBC5',
                "0.0151 - 0.025" = '#FFFFCC')

# Quantile colour palette - blue-based
color_palette <- c("Below 0" = '#FFFFCC', 
                "0 - 0.0072" = '#DEEBF7',
                "0.0072 - 0.0104" = '#C6DBEF',
                "0.0104 - 0.0151" = '#6BAED6',
                "0.0151 - 0.025" = '#2171B5')

plot <- ggplot(data = merged_data2) +
  geom_sf(aes(fill = coef_li_exp_cat), color = NA) +  # 'color = NA' removes polygon borders
  scale_fill_manual(values = color_palette, 
                    name = expression(italic(LI)[k] ~ "Coefficient"),
                    drop = FALSE) +
  # Remove background, grid, axis labels, and ticks
  theme_void() +

  theme(
    panel.background = element_rect(fill = 'white', color = NA),
    legend.position = c(0.99, 0.03),  
    legend.justification = c("right", "bottom"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )+

  # Reduce margins
  theme(plot.margin = margin(0, 0, 0, 0))

plot

ggsave('C:/Users/phily/liveability-index/output/active travel/gwr_coef.png', plot, dpi = 300, width = 8, height = 6)
```


```{r gwr pseudo t-statistics}
# Pseudo t-statistics
# Extract coefficients and standard errors
coefs <- gwr.model$SDF$"li_exp"
ses   <- gwr.model$SDF$"li_exp_se"

# coefs_1 <- gwr.model$SDF$"(Intercept)"
# ses_1 <- gwr.model$SDF$"(Intercept)"

# Compute pseudo t-values
pseudo_tvals <- coefs / ses %>% as_tibble() %>% rename(pseudo_t = value)


# 2. Add classification by significance level
pseudo_tvals <- pseudo_tvals %>%
  mutate(
    significance = case_when(
      abs(pseudo_t) > 2.58 ~ "99% (|t| > 2.58)",
      abs(pseudo_t) > 1.96 ~ "95% (|t| > 1.96)",
      TRUE ~ "Not significant"
    )
  )




#store significance results
merged_data2 <- cbind(merged_data1, pseudo_tvals)

# 4. Plot using ggplot2 (optional)
# If you want to visualize with coordinates:
library(ggplot2)



plot2 <- ggplot(merged_data2) +
  geom_sf(aes(fill = significance), color = "white", size = 0.2) +
  scale_fill_manual(
    values = c(
      "99% (|t| > 2.58)" = '#B2E2B4', 
      "95% (|t| > 1.96)" = '#FFFFCC',
      "Not significant" = 'grey70'
    ),
    labels = c(
      "99% (|t| > 2.58)" = expression("99% (" * "|" * italic(t) * "|" * " > 2.58)"),
      "95% (|t| > 1.96)" = expression("95% (" * "|" * italic(t) * "|" * " > 1.96)"),
      "Not significant" = "Not significant"
    )
  ) +
  theme_void() +
  labs(fill = "Significance Level") +
  theme(
    panel.background = element_rect(fill = 'white', color = NA),
    legend.position = c(0.99, 0.03),  
    legend.justification = c("right", "bottom"),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

plot2

ggsave('C:/Users/phily/liveability-index/output/active travel/gwr_coef_sig.png', plot2, dpi = 300, width = 8, height = 6)
```

The GWR results show that liveability (li_exp) is positively associated with footfall in most neighbourhoods across London. However, the magnitude of this relationship varies geographically. Notably, Outer and Southeast London exhibit stronger local coefficients, suggesting that improvements in liveability are more strongly tied to increased footfall in these areas. This is particularly insightful for urban planning for low liveability neighbourhoods in Outer London as investment in improving neighbourhood profile could significantly bolster footfall patterns in these areas more than other areas.

The statistical significance test confirms that these coefficients are significantly different from zero in a large portion of both Inner and Outer London, reinforcing the robustness of the observed spatial variation.


```{r residuals vs fitted value}
# Extract residuals
residuals <- gwr.model$SDF$gwr.e

# Alternatively, also get fitted values if needed
fitted_vals <- gwr.model$SDF$pred

# Assuming fitted_vals and residuals are vectors of your fitted values and residuals
df <- data.frame(Fitted = fitted_vals, Residuals = residuals)

plot <- ggplot(df, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.6, color = "#2c7fb8", size = 2) +      # semi-transparent blue points
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  # horizontal line at zero
  theme_minimal(base_size = 14) +                            # clean minimal theme
  labs(title = "Residuals vs Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

ggsave('C:/Users/phily/liveability-index/output/active travel/gwr_residuals.png', plot, dpi = 300, width = 8, height = 6)

# Moran's I test on GWR residuals
# save residuals of gwr
merged_data1 <- merged_data1 %>%
  mutate(gwr_res = residuals)

Nearest_neighbour <- merged_data1 %>%
  st_drop_geometry()%>%
  dplyr::select(gwr_res)%>%
  pull()%>%
  moran.test(., LNeigh.knn_4_weight)

Nearest_neighbour
```


